---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  release: "Unknown"
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  fisher_1per_timeframe: "Annual"
  fisher_1per_year_selection: "None"
  fisher_1per_period_selection: "None"
  fisher_1per_lobster_season_selection: "None" 
  fisher_1per_conch_season_selection: "None" 
  fisher_1per_finfish_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Determine year or season  ---------------------------
timeframe <- "Annual"

# Filter to year or season  ---------------------------
df <- params$datafile[[1]]
df$`Waypoint Date` <- as.Date(df$`Waypoint Date`)
if (timeframe == "Annual" && as.character(params$fisher_1per_year_selection) != "None") {
  df <- df |>
    filter(format(`Waypoint Date`, "%Y") == params$fisher_1per_year_selection)
}

# Add month column  ---------------------------
df$Month <- format(df$`Waypoint Date`, "%b %Y")

# Define study timeframe ---------------------------
study_start <- min(df$`Waypoint Date`, na.rm = TRUE)
study_end <- max(df$`Waypoint Date`, na.rm = TRUE)
study_timeframe <- unique(format(df$`Waypoint Date`, "%Y"))[1]
```

# Turneffe Atoll Fisher Catch Finfish `r timeframe` Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)


***


The marine resources and biodiversity abundant in coral reef ecosystems support the 
livelihoods of coastal communities, and provide economic opportunities that can be 
sustainable given attentive management and enforcement. 
Monitoring the activities and harvests of local fishers is important for 
understanding ecosystem productivity, assessing stock sustainability, 
and informing effective management decisions. 
This report presents `r tolower(timeframe)` findings from fishery-dependent monitoring of finfish 
at Turneffe Atoll Marine Reserve (TAMR) from `r study_timeframe`. 
Data were recorded directly by fishermen using SMART technology through a data collection 
package designed by TASA. This report aims to summarize and visualize the Fisher 
Finfish Catch dataset from `r study_timeframe` to facilitate data interpretation and support 
management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Fisher Catch Project

The Fisher Catch Data Program is a pilot initiative led by the Turneffe Atoll Sustainability 
Association (TASA) and funded by the Inter-American Development Bank (IDB). The 
program aims to evaluate the effectiveness of using digital technology for catch 
reporting by fishers in the General Use Zone of the Turneffe Atoll Marine Reserve 
(TAMR), specifically for Queen Conch, Caribbean Spiny Lobster, and finfish. By 
transitioning from paper-based data sheets to a customized, simplified version of the 
SMART Fisheries Catch Data module, the program streamlines data collection and 
submission. Fishers were equipped with a complete kit—including a SMART-enabled 
device, measuring tools, and training materials—and participated in a three-day training 
session on standardized data collection and entry. This fishery-dependent monitoring 
effort gathers key biological and spatial data such as species composition, size, gear 
type, fishing effort, and location. The information collected will help assess fishing 
pressure, track Catch per Unit Effort (CPUE), ultimately guiding adaptive fisheries 
management strategies.

The primary objective of the Fisher Catch Program at the Turneffe Atoll Marine Reserve 
(TAMR) is to assess whether fishers can effectively use digital technology to report catch 
data on Queen Conch, Caribbean Spiny Lobster, and finfish. By replacing traditional 
paper-based methods, the program aims to simplify and standardize data collection and 
submission, while generating accurate, fishery-dependent data to inform adaptive 
fisheries management, track trends in fishing pressure and landings.\

### Finfish

Finfish refer to all fish species that can be classified as ‘true fish,’ in contrast to other aquatic animals 
such as shellfish or cuttlefish.
In Belize, over 650 species of finfish may be found (FishBase 2024), with many of these, such as 
Mutton Snapper (_Lutjanus analis_),
Nassau Grouper (_Epinephelus striatus_), and Queen Triggerfish (_Balistes vetula_) being 
listed as Vulnerable conservation status on the IUCN.

Finfish provide a variety of services to coral reefs, including grazing of macroalgae and predation.
Fishing for finfish may be done using handline fishing or winch (targeting snappers, jacks, groupers, and grunts primarily),
long lines (targeting pelagic species and sharks), or gillnets (targeting sharks) in Belize (UNCTAD 2020).
Finfish fishing in Belize is regulated by the Belize Fisheries Department (BFD),
with both general regulations such as fishers requiring fishing licenses and color-coded vessels, and specific regulations,
such as forbidding the fishing of parrotfish and tangs (Scaridae) (Belize Fisheries Department accessed 2025-b).
To maintain healthy stocks and ecological services of finfish, it is important to maintain the size, reproductive capacity,
and biodiversity of key taxa.\


## Methods

Fisher catch data were collected at Turneffe Atoll 
Marine Reserve (TAMR) (Figure 1). 
The survey period for this report’s dataset extended 
from `r study_start` to `r study_end` at TAMR.

The fishermen participating in the Fisher Catch Program at TAMR were selected based 
on their active use of the General Use Zone within the Turneffe Atoll Marine Reserve 
and their willingness to participate in a pilot program focused on digital data collection. 
TASA prioritized individuals who represented the main fishing groups in the 
area—sailboat fishers, skiff fishers, and camp fishers—to ensure broad coverage of 
fishing practices and effort across different habitats. Selection also considered the 
fishers’ availability for training and their interest in adopting technology to support 
sustainable fisheries management.

Data on gutted or whole weight (lbs), total length (Inches), and fork length (Inches) were entered by fishermen using SMART 
technology through a data collection package 
designed by TASA. 
The process of entering data by fishermen into SMART in the Fisher Catch Program follows a 
structured and simplified workflow designed for ease of use in the field:

1. **Catch and Effort Recording**: After each fishing trip, the fisher records details of the 
   catch, including species (Queen Conch, Caribbean Spiny Lobster, or finfish), quantity, 
   size (measured using calipers, scales, or fishboards), gear type used, hours fished, and 
   fishing location. 

2. **Use of SMART Device**: The fisher uses the provided SMART-enabled device (typically a 
   smartphone) to input the recorded data directly into the customized, user-friendly 
   SMART Fisheries Catch Data module. 

3. **Guided Input**: The module includes drop-down menus, visual aids, and pre-set data 
   fields to guide fishers through the entry process, reducing errors and ensuring 
   consistency. 

4. **Spatial Data Entry**: Spatial information (coordinates) are automatically recorded 
   showing where fishing took place. 

5. **Review and Save**: Before finalizing, fishers can review their entries for accuracy. Once 
   confirmed, the data is saved locally on the device. 

6. **Data Sync**: When the device connects to the internet (either at sea if possible or upon 
   return to shore), the data is synced with the central SMART database managed by TASA 
   for validation, analysis, and reporting.


This report, including text, figures, and tables, was generated using the 
Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025, version `r params$release`), an R Shiny app which takes 
standardized reef monitoring data, validates the dataset, and creates a shareable output. \


```{r figure1, echo = FALSE, fig.width = 7, fig.height = 7}
# Create map for Figure 1 ---------------------------
df_map <- df |>
  select(`Waypoint ID`, Y, X) |>
  distinct()
if (length(unique(df$`Waypoint ID`)) < 5) {
  generate_map(df_map, "Y", "X", "Waypoint ID")
} else {
  generate_map_nolabels(df_map, "Y", "X", "Waypoint ID")
}
```


*Figure 1.* Map of Turneffe Atoll Marine Reserve zones and ecosystems. 
Shapefiles for zones belong to the Belize Fisheries Department (Belize Geomatics 2022). 
Shapefiles for the base map and ecosystems come from the publicly available Spatial Data 
Warehouse for Belize (Meerman 2013, Meerman 2017).
Waypoints for finfish recorded in `r study_timeframe` are labelled on the map.\

 
## Results


```{r table1_prep, echo = FALSE, include = FALSE}
df_events_overall <- df %>%
  select(`Waypoint ID`, `Hours Fished`, `Total Lbs of Catch`) %>%
  distinct() %>%
  filter(!is.na(`Hours Fished`), !is.na(`Total Lbs of Catch`), `Hours Fished` > 0) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE),
    `Catch (lb)` = sum(`Total Lbs of Catch`, na.rm = TRUE)
  ) %>%
  mutate(Month = "Overall")
df_table1_overall <- df %>%
  summarize(
    Fishers = n_distinct(`Fishers Name`),
    Finfish = n(),
    Events = n_distinct(`Waypoint ID`)
  ) %>%
  mutate(Month = "Overall") %>%
  left_join(df_events_overall, by = "Month") %>%
  mutate(
    `CPUE (lb/hr)` = ifelse(Hours > 0, round(`Catch (lb)` / Hours, 1), NA),
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  )
df_events <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  select(Month, `Waypoint ID`, `Hours Fished`, `Total Lbs of Catch`) %>%
  distinct() %>%
  filter(!is.na(`Hours Fished`), !is.na(`Total Lbs of Catch`), `Hours Fished` > 0) %>%
  group_by(Month) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE),
    `Catch (lb)` = sum(`Total Lbs of Catch`, na.rm = TRUE)
  )
df_table1 <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Fishers = n_distinct(`Fishers Name`),
    Finfish = n(),
    Events = n_distinct(`Waypoint ID`)
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  mutate(
    `CPUE (lb/hr)` = ifelse(Hours > 0, round(`Catch (lb)` / Hours, 1), NA),
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  ) %>%
  bind_rows(df_table1_overall)

# Prepare dynamic variables for Table 1 ---------------------------
total_fishers <- length(unique(df$`Fishers Name`))
total_finfish <- df_table1_overall$Finfish
total_events <- df_table1_overall$Events
total_hours <- df_table1_overall$Hours
max_cpue_lb <- max(df_table1$`CPUE (lb/hr)`, na.rm = TRUE)
max_cpue_lb_month <- df_table1$Month[which.max(df_table1$`CPUE (lb/hr)`)]
min_cpue_lb <- min(df_table1$`CPUE (lb/hr)`, na.rm = TRUE)
min_cpue_lb_month <- df_table1$Month[which.min(df_table1$`CPUE (lb/hr)`)]
max_cpue_num <- max(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
max_cpue_num_month <- df_table1$Month[which.max(df_table1$`CPUE (num/hr)`)]
min_cpue_num <- min(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
min_cpue_num_month <- df_table1$Month[which.min(df_table1$`CPUE (num/hr)`)]
```


Between `r study_start` and `r study_end`, `r total_fishers` fishermen contributed data to the queen finfish SMART dataset. 
A total of `r total_finfish` finfish were caught across `r total_events` 
fishing events, totalling `r total_hours` active fishing hours. 
The highest catch per unit effort (CPUE) 
in pounds per hour (lb/hr) occurred in `r max_cpue_lb_month` with `r max_cpue_lb`, whereas the lowest 
occurred in `r min_cpue_lb_month` with `r min_cpue_lb`. The highest CPUE in count 
per hour (num/hr) occurred in `r max_cpue_num_month` 
with `r max_cpue_num`, whereas the lowest was `r min_cpue_num_month` 
with `r min_cpue_num` (Table 1; Figure 2).\


**Table 1.** Summary of fishing activity by month, including total fishers, events, and finfish caught, 
along with fishing hours, pounds of catch, and catch per unit effort (CPUE). 
CPUE values are calculated only for events where fishing hours and catch weight were recorded. 
Data were collected using SMART finfish data from fishers at Turneffe Atoll Marine Reserve, `r study_timeframe`.\

```{r table1, echo = FALSE}
# show table
as.data.frame(df_table1)
```


<br>


```{r table2_prep, echo = FALSE, include = FALSE}
# Create df for Table 2 ---------------------------
df_table2_rows <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  group_by(`Waypoint ID`) %>%
  summarize(
    Month = first(Month),
    Month_sort = first(Month_sort),
    Hours = first(`Hours Fished`),
    Catch_lb = first(`Total Lbs of Catch`),
    Finfish = n(),
    .groups = "drop"
  ) %>%
  mutate(
    `CPUE (lb/hr)` = ifelse(Hours > 0, Catch_lb / Hours, NA),
    `CPUE (num/hr)` = ifelse(Hours > 0, Finfish / Hours, NA)
  )
df_table2_lb_overall <- df_table2_rows %>%
  summarize(
    Min = round(min(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (lb/hr)`, na.rm = TRUE), 1)
  ) %>%
  mutate(Month = "Overall", CPUE = "lb/hr")
df_table2_lb <- df_table2_rows %>%
  mutate(CPUE = "lb/hr") %>%
  group_by(CPUE, Month, Month_sort) %>%
  summarize(
    Min = round(min(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (lb/hr)`, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  bind_rows(df_table2_lb_overall)
df_table2_num_overall <- df_table2_rows %>%
  summarize(
    Min = round(min(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (num/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (num/hr)`, na.rm = TRUE), 1)
  ) %>%
  mutate(Month = "Overall", CPUE = "num/hr")
df_table2_num <- df_table2_rows %>%
  mutate(CPUE = "num/hr") %>%
  group_by(CPUE, Month, Month_sort) %>%
  summarize(
    Min = round(min(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (num/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (num/hr)`, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  bind_rows(df_table2_num_overall)
df_table2 <- df_table2_lb_overall %>%
  bind_rows(df_table2_num_overall) %>%
  select(CPUE, Min, Mean, Median, Max, SD)


# Prepare dynamic variables for Table 2 ---------------------------
overall_mean <- df_table2_lb_overall$Mean
overall_median <- df_table2_lb_overall$Median
overall_mean_num <- df_table2_num_overall$Mean
overall_median_num <- df_table2_num_overall$Median
max_cpue_mean_lb <- max(df_table2_lb$Mean, na.rm = TRUE)
max_cpue_mean_lb_month <- df_table2_lb$Month[which.max(df_table2_lb$Mean)]
min_cpue_mean_lb <- min(df_table2_lb$Mean, na.rm = TRUE)
min_cpue_mean_lb_month <- df_table2_lb$Month[which.min(df_table2_lb$Mean)]
max_cpue_mean_num <- max(df_table2_num$Mean, na.rm = TRUE)
max_cpue_mean_num_month <- df_table2_num$Month[which.max(df_table2_num$Mean)]
min_cpue_mean_num <- min(df_table2_num$Mean, na.rm = TRUE)
min_cpue_mean_num_month <- df_table2_num$Month[which.min(df_table2_num$Mean)]
```


The CPUE of finfish in pounds per hour had a 
mean of `r overall_mean`lb/hr and median of `r overall_median`lb/hr, 
with the highest mean of `r max_cpue_mean_lb`lb/hr being in `r max_cpue_mean_lb_month` 
and the lowest mean of `r min_cpue_mean_lb`lb/hr being in `r min_cpue_mean_lb_month`. 
The CPUE of finfish in count per hour had a 
mean of `r overall_mean_num`num/hr and median of `r overall_median_num`num/hr, 
with the highest mean of `r max_cpue_mean_num`num/hr being in `r max_cpue_mean_num_month` 
and the lowest mean of `r min_cpue_mean_num`num/hr being in `r min_cpue_mean_num_month`. 
(Table 2; Figure 2).\




```{r figure2_prep, echo = FALSE, include = FALSE}
# Create dfs for Figure 2 ---------------------------
df_fig2_A <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Fishers = n_distinct(`Fishers Name`),
    Finfish = n(),
    Events = n_distinct(`Waypoint ID`),
    .groups = "drop"
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  mutate(
    `CPUE (lb/hr)` = ifelse(Hours > 0, round(`Catch (lb)` / Hours, 1), NA),
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  )

# Create panels for Figure 2 ---------------------------
A <- ggplot(df_fig2_A, aes(x = Month_sort, y = Events, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Fishing Events") +
  labs(title = "A")
B <- ggplot(df_fig2_A, aes(x = Month_sort, y = Hours, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Hours Fished") +
  labs(title = "B")
C <- ggplot(df_fig2_A, aes(x = Month_sort, y = Finfish, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Number Finfishs") +
  labs(title = "C")
D <- ggplot(df_fig2_A, aes(x = Month_sort, y = `Catch (lb)`, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Catch (lb)") +
  labs(title = "D")
E <- ggplot(df_fig2_A, aes(x = Month_sort, group = 1)) +
  geom_line(aes(y = `CPUE (lb/hr)`), size = 0.5) +
  geom_point(aes(y = `CPUE (lb/hr)`, shape = "lb/hr"), size = 2.5) +
  geom_line(aes(y = `CPUE (num/hr)`), size = 0.5) +
  geom_point(aes(y = `CPUE (num/hr)`, shape = "num/hr"), size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  scale_shape_manual(
    name = "CPUE Type",
    values = c("lb/hr" = 4, "num/hr" = 19)
  ) +
  theme_custom +
  xlab("Month") +
  ylab("Catch per Unit Effort") +
  labs(title = "E")


# Assemble Figure 2 ---------------------------
figure_2_top <- ggarrange(
  A, B, C, D,
  ncol = 2, nrow = 2
)
figure2 <- ggarrange(
  figure_2_top, E,
  ncol = 1,
  heights = c(2, 1)
)
```


```{r figure2, fig.width=6, fig.height=8, fig.align = 'left', echo = FALSE}
figure2
```


*Figure 2.* Line plots of fishing events (A), 
active hours fished (B), finfish caught (C), 
and catch per unit effort (in pounds and count per hour) (D) 
over time, collected using SMART finfish data from fishers at Turneffe 
Atoll Marine Reserve, `r study_timeframe`.\


**Table 2.** Summary statistics for catch per unit effort in lb/hr and count/hr, 
including minimum, mean, median, and maximum values, 
along with standard deviation. 
CPUE values are calculated only for events where fishing hours and catch weight were recorded. 
Data were collected using SMART finfish data from fishers at Turneffe Atoll Marine Reserve, `r study_timeframe`.\
```{r table2, echo = FALSE}
# show table
df_table2
```


<br>


```{r table3_prep, echo = FALSE, include = FALSE}
# Create df for Table 3 ---------------------------
df_table3 <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    `Mean Length (In)` = round(mean(as.numeric(`Total Length (Inches)`), na.rm = TRUE), 1),
    `Median Length (In)` = round(median(as.numeric(`Total Length (Inches)`), na.rm = TRUE), 1),
    `Mean Fork Length (In)` = round(mean(as.numeric(`Fork Length (Inches)`), na.rm = TRUE), 1),
    `Median Fork Length (In)` = round(median(as.numeric(`Fork Length (Inches)`), na.rm = TRUE), 1),
    `Mean Weight Gutted (lbs)` = ifelse(any(`Weight Type` == "Gutted"),
      round(mean(as.numeric(`Weight (lbs)`)[`Weight Type` == "Gutted"], na.rm = TRUE), 1),
      NA
    ),
    `Median Weight Gutted (lbs)` = ifelse(any(`Weight Type` == "Gutted"),
      round(median(as.numeric(`Weight (lbs)`)[`Weight Type` == "Gutted"], na.rm = TRUE), 1),
      NA
    ),
    `Mean Weight Whole (lbs)` = ifelse(any(`Weight Type` == "Whole"),
      round(mean(as.numeric(`Weight (lbs)`)[`Weight Type` == "Whole"], na.rm = TRUE), 1),
      NA
    ),
    `Median Weight Whole (lbs)` = ifelse(any(`Weight Type` == "Whole"),
      round(median(as.numeric(`Weight (lbs)`)[`Weight Type` == "Whole"], na.rm = TRUE), 1),
      NA
    )
  ) %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  bind_rows(
    df %>%
      summarize(
        `Mean Length (In)` = round(mean(as.numeric(`Total Length (Inches)`), na.rm = TRUE), 1),
        `Median Length (In)` = round(median(as.numeric(`Total Length (Inches)`), na.rm = TRUE), 1),
        `Mean Fork Length (In)` = round(mean(as.numeric(`Fork Length (Inches)`), na.rm = TRUE), 1),
        `Median Fork Length (In)` = round(median(as.numeric(`Fork Length (Inches)`), na.rm = TRUE), 1),
        `Mean Weight Gutted (lbs)` = ifelse(any(`Weight Type` == "Gutted"),
          round(mean(as.numeric(`Weight (lbs)`)[`Weight Type` == "Gutted"], na.rm = TRUE), 1),
          NA
        ),
        `Median Weight Gutted (lbs)` = ifelse(any(`Weight Type` == "Gutted"),
          round(median(as.numeric(`Weight (lbs)`)[`Weight Type` == "Gutted"], na.rm = TRUE), 1),
          NA
        ),
        `Mean Weight Whole (lbs)` = ifelse(any(`Weight Type` == "Whole"),
          round(mean(as.numeric(`Weight (lbs)`)[`Weight Type` == "Whole"], na.rm = TRUE), 1),
          NA
        ),
        `Median Weight Whole (lbs)` = ifelse(any(`Weight Type` == "Whole"),
          round(median(as.numeric(`Weight (lbs)`)[`Weight Type` == "Whole"], na.rm = TRUE), 1),
          NA
        )
      ) %>%
      mutate(Month = "Overall")
  ) %>%
  mutate(across(everything(), ~ {
    if (is.numeric(.)) {
      ifelse(is.na(.) | is.nan(.), "-", as.character(.))
    } else {
      ifelse(is.na(.), "-", .)
    }
  }))

# Prepare dynamic variables for Table 3 ---------------------------
total_weight_gutted <- filter(df_table3, Month == "Overall")$`Mean Weight Gutted (lbs)`
max_weight_gutted <- max(filter(df_table3, Month != "Overall")$`Mean Weight Gutted (lbs)`, na.rm = TRUE)
max_weight_gutted_month <- df_table3$Month[which.max(filter(df_table3, Month != "Overall")$`Mean Weight Gutted (lbs)`)]
total_weight_whole <- filter(df_table3, Month == "Overall")$`Mean Weight Whole (lbs)`
max_weight_whole <- max(filter(df_table3, Month != "Overall")$`Mean Weight Whole (lbs)`, na.rm = TRUE)
max_weight_whole_month <- df_table3$Month[which.max(filter(df_table3, Month != "Overall")$`Mean Weight Whole (lbs)`)]
total_total_length <- filter(df_table3, Month == "Overall")$`Mean Length (In)`
max_total_total_length <- max(filter(df_table3, Month != "Overall")$`Mean Length (In)`, na.rm = TRUE)
max_total_total_length_month <- df_table3$Month[which.max(filter(df_table3, Month != "Overall")$`Mean Length (In)`)]
total_fork_length <- filter(df_table3, Month == "Overall")$`Mean Fork Length (In)`
max_total_fork_length <- max(filter(df_table3, Month != "Overall")$`Mean Fork Length (In)`, na.rm = TRUE)
max_total_fork_length_month <- df_table3$Month[which.max(filter(df_table3, Month != "Overall")$`Mean Fork Length (In)`)]
```

Average total length was `r total_total_length`in, greatest in `r max_total_total_length_month` with `r max_total_total_length`in. 
Average fork length was `r total_fork_length`in, greatest in `r max_total_fork_length_month` with `r max_total_fork_length`in. 
Average gutted weight was `r total_weight_gutted`lbs, greatest in `r max_weight_gutted_month` with `r max_weight_gutted`lbs. 
Average whole weight was `r total_weight_whole`lbs, greatest in `r max_weight_whole_month` with `r max_weight_whole`lbs 
(Table 3; Figure 3).\


**Table 3.** Summary of mean and median total length (Inches), fork length (Inches), gutted weight (lbs), and whole weight(lbs) of finfish, 
using SMART finfish data from fishers at Turneffe Atoll Marine Reserve, `r study_timeframe`.

```{r table3, echo = FALSE}
# show table
as.data.frame(df_table3)
```


<br>


```{r figure3_prep, echo = FALSE, include = FALSE}
# Create dfs for Figure 3 ---------------------------
gear_levels <- df %>%
  pull(`Type of Gear`) %>%
  unique() %>%
  sort()
gear_colors <- setNames(palette[2:(1 + length(gear_levels))], gear_levels)
df_fig3_unsummarized <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01")),
    weight_for_gutted = ifelse(`Weight Type` == "Gutted", `Weight (lbs)`, NA_real_),
    weight_for_whole = ifelse(`Weight Type` == "Whole", `Weight (lbs)`, NA_real_)
  ) %>%
  arrange(Month_sort)
df_fig3 <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01")),
    weight_for_gutted = ifelse(`Weight Type` == "Gutted", `Weight (lbs)`, NA_real_),
    weight_for_whole = ifelse(`Weight Type` == "Whole", `Weight (lbs)`, NA_real_)
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    `Mean Length (In)` = round(mean(as.numeric(`Total Length (Inches)`), na.rm = TRUE), 1),
    `Mean Fork Length (In)` = round(mean(as.numeric(`Fork Length (Inches)`), na.rm = TRUE), 1),
    `Mean Weight Gutted (lbs)` = ifelse(any(`Weight Type` == "Gutted"),
      round(mean(as.numeric(`Weight (lbs)`)[`Weight Type` == "Gutted"], na.rm = TRUE), 1),
      NA
    ),
    `Mean Weight Whole (lbs)` = ifelse(any(`Weight Type` == "Whole"),
      round(mean(as.numeric(`Weight (lbs)`)[`Weight Type` == "Whole"], na.rm = TRUE), 1),
      NA
    )
  ) %>%
  arrange(Month_sort)
df_fig3_gutted_weight <- df_fig3 %>%
  filter(!is.nan(`Mean Weight Gutted (lbs)`))
df_fig3_whole_weight <- df_fig3 %>%
  filter(!is.nan(`Mean Weight Whole (lbs)`))
df_fig3_total_length <- df_fig3 %>%
  filter(!is.nan(`Mean Length (In)`))
df_fig3_fork_length <- df_fig3 %>%
  filter(!is.nan(`Mean Fork Length (In)`))

# Create panels for Figure 3 ---------------------------
A <- ggplot(df_fig3_unsummarized, aes(x = `Type of Gear`, fill = `Type of Gear`)) +
  geom_bar(color = "black") +
  scale_fill_manual(values = gear_colors, drop = FALSE) +
  theme_custom +
  expand_limits(x = 0) +
  xlab("") +
  ylab("Count") +
  labs(title = "A") +
  theme(axis.text.x = element_text(angle = 0))
B <- ggplot(df_fig3_unsummarized, aes(x = `Total Length (Inches)`, fill = `Type of Gear`)) +
  geom_histogram(binwidth = 1, color = "black") +
  scale_fill_manual(values = gear_colors, drop = FALSE) +
  theme_custom +
  expand_limits(x = 0) +
  xlab("Total Length (Inches)") +
  ylab("Count") +
  labs(title = "B") +
  theme(legend.position = "none")
C <- ggplot() +
  geom_line(
    data = df_fig3_total_length,
    aes(x = Month_sort, y = `Mean Length (In)`, linetype = "Total Length"),
    size = 0.5
  ) +
  geom_point(
    data = df_fig3_total_length,
    aes(x = Month_sort, y = `Mean Length (In)`, shape = "Total Length"),
    size = 2.5
  ) +
  geom_line(
    data = df_fig3_fork_length,
    aes(x = Month_sort, y = `Mean Fork Length (In)`, linetype = "Fork Length"),
    size = 0.5
  ) +
  geom_point(
    data = df_fig3_fork_length,
    aes(x = Month_sort, y = `Mean Fork Length (In)`, shape = "Fork Length"),
    size = 2.5
  ) +
  theme_custom +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  ylab("Mean Total Length (Inches)") +
  xlab("Month") +
  labs(
    title = "C",
    linetype = "",
    shape = ""
  ) +
  scale_linetype_manual(
    values = c("Total Length" = "solid", "Fork Length" = "dotted")
  ) +
  scale_shape_manual(
    values = c("Total Length" = 16, "Fork Length" = 4)
  )
D <- ggplot(df_fig3_gutted_weight, aes(x = Month_sort, y = `Mean Weight Gutted (lbs)`)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  theme_custom +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  ylab("Gutted Weight (lb)") +
  xlab("Month") +
  labs(title = "D")
E <- ggplot(df_fig3_whole_weight, aes(x = Month_sort, y = `Mean Weight Whole (lbs)`)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  theme_custom +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  ylab("Whole Weight (lb)") +
  xlab("Month") +
  labs(title = "E")

# Assemble Figure 3 ---------------------------
figure_3_bottom <- ggarrange(
  B, C, D, E,
  ncol = 2, nrow = 2
)
figure3 <- ggarrange(
  A, figure_3_bottom,
  ncol = 1,
  heights = c(1, 2.4)
)
```


```{r figure3, fig.width=6, fig.height=8, fig.align = 'left', echo = FALSE}
figure3
```


*Figure 3.* Bar plot of fishing gear types (A), 
histogram of fish length (Inches) by gear type (B), 
line plot of mean fish total and fork length (Inches) over time (C), 
line plot of gutted weight (lbs) over time (D), and line plot of whole weight (lbs) over time (E), 
collected using SMART finfish data from fishers at Turneffe 
Atoll Marine Reserve, `r study_timeframe`.\


```{r figure4_prep, echo = FALSE, include = FALSE}
# Create df for Figure 4 ---------------------------
df_figure4_totals <- df %>%
  group_by(`Fish Species`) %>%
  summarise(`Total Count` = n(), ..groups = "drop")
df_figure4 <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  group_by(`Fish Species`, Month, Month_sort) %>%
  summarise(Count = n()) %>%
  left_join(df_figure4_totals, by = "Fish Species") %>%
  mutate(`Fish Species` = as.factor(`Fish Species`))

# Create Figure 4 ---------------------------
figure4 <- ggplot(df_figure4, aes(
  x = fct_reorder(`Fish Species`, desc(`Total Count`)),
  y = Count,
  fill = fct_reorder(Month, Month_sort)
)) +
  geom_col(color = "black") +
  scale_fill_manual(values = palette_extended) +
  theme_custom +
  expand_limits(y = 0) +
  xlab("Fish Species") +
  ylab("Count") +
  labs(fill = "") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# Prepare dynamic variables for Figure 4 ---------------------------
common_fish <- arrange(df_figure4_totals, desc(`Total Count`))$`Fish Species`
common_fish_counts <- arrange(df_figure4_totals, desc(`Total Count`))$`Total Count`
unique_months <- unique(arrange(df_figure4, Month_sort)$Month)
common_fish_text <- c()
for (month in unique_months) {
  common_fish_text <- c(
    common_fish_text, paste0(
      "the most common fish in ",
      month,
      " was ",
      arrange(filter(df_figure4, Month == month), desc(Count))$`Fish Species`[1]
    )
  )
}
common_fish_text <- knitr::combine_words(common_fish_text)
```


The most common fish caught in `r study_timeframe` were 
`r common_fish[1]` (`r common_fish_counts[1]` individuals), 
`r common_fish[2]` (`r common_fish_counts[2]` individuals), and 
`r common_fish[3]` (`r common_fish_counts[3]` individuals). 
Across months, `r common_fish_text` (Figure 4).\


```{r figure4, fig.width=6, fig.height=6, fig.align = 'left', echo = FALSE}
figure4
```


*Figure 4.* Column plot of commonly caught fish species by month in which they were caught, 
collected using SMART finfish data from fishers at Turneffe 
Atoll Marine Reserve, `r study_timeframe`.\


<br>


## Conclusion

By summarizing and visualizing the fisher catch finfish SMART data from `r study_timeframe`, 
researchers, policy-makers, and other stakeholders are better able to interpret and use the dataset. 
This report supports the aims of the Fisher Catch SMART data collection by offering a more understandable view 
of the data itself, supporting future efforts to create recommendations on enforcement, monitoring, 
education, outreach, and regulations to ensure the long-term sustainability of the TAMR fishery while safeguarding the 
ecological value of the atoll.\


## Acknowledgments

Thank you to the fishers who contributed to this project.\


## Report Metadata

**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\
**App Version:** `r params$release`\


## Literature Cited

Belize Fisheries Department. (n.d.). Fisheries Regulations. https://fisheries.gov.bz/regulations/

Fedler, A. J. (2011). The Economic Value of Turneffe Atoll (Full Report). Turneffe Atoll Trust. http://www.turneffeatollmarinereserve.org/app/webroot/userfiles/214/File/Turneffe%20Atoll%20Valuation%20Study%20PDF%20-%20Final%20Report.pdf

Froese, R., & Pauly, D. (2024). FishBase [World Wide Web electronic publication.]. www.fishbase.org

Meerman, J. C. (2013). Belize Basemap [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Meerman, J. C. (2017). Belize_Ecosystems_2017 [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Stoddart, D. R. (1962). Three Caribbean atolls: Turneffe Islands, Lighthouse Reef and Glover’s Reef,  British Honduras. Atoll Research Bulletin, 87, 1–151.

Turneffe Atoll Sustainability Association. (2025). Turneffe Reef-Monitoring Data Reporting Tool (Version 1.0) [Computer software].

United Nations Conference (UNCTAD) & on Trade and Development. (2020). Oceans Economy and Trade Strategy: Belize Marine Fisheries and Seafood Processing (UNCTAD/DITC/TED/INF/2020/5). https://unctad.org/system/files/official-document/ditctedinf2020d5_en.pdf