---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  release: "Unknown"
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  fisheries_1per_timeframe: "Annual"
  fisheries_1per_year_selection: "None"
  fisheries_1per_period_selection: "None"
  fisheries_1per_lobster_season_selection: "None" 
  fisheries_1per_conch_season_selection: "None" 
  fisheries_1per_finfish_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Determine year or season  ---------------------------
timeframe <- params$fisheries_1per_timeframe

# Filter to year or season  ---------------------------
df <- params$datafile[[1]]
df$`Date of Encounter` <- as.Date(df$`Date of Encounter`)
if (timeframe == "Annual" && as.character(params$fisheries_1per_year_selection) != "None") {
  df <- df |>
    filter(format(`Date of Encounter`, "%Y") == params$fisheries_1per_year_selection)
} else if (timeframe == "Seasonal" && as.character(params$fisheries_1per_conch_season_selection) != "None") {
  season_years <- unlist(strsplit(params$fisheries_1per_conch_season_selection, "-"))
  start_date <- as.Date(paste0(season_years[1], "-07-01"))
  end_date <- as.Date(paste0(season_years[2], "-02-28"))
  df <- df |>
    filter(`Date of Encounter` >= start_date, `Date of Encounter` <= end_date)
}

# Add month column  ---------------------------
df$Month <- format(df$`Date of Encounter`, "%b %Y")

# Define study timeframe ---------------------------
study_start <- min(df$`Date of Encounter`, na.rm = TRUE)
study_end <- max(df$`Date of Encounter`, na.rm = TRUE)
if (timeframe == "Annual") {
  study_timeframe <- unique(format(df$`Date of Encounter`, "%Y"))[1]
} else if (timeframe == "Seasonal") {
  study_timeframe <- paste0(format(study_start, "%Y"), "-", format(study_end, "%Y"))
}
```

# Turneffe Atoll Fisheries Catch Conch `r timeframe` Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)


**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\
**App Version:** `r params$release`\

***

Coral reef ecosystems, like those of Turneffe Atoll Marine Reserve (TAMR), provide key ecological and economic value to coastal families. 
With proper management and monitoring, 
these ecosystems can continue to support people sustainably in the long-term. 
Directly tracking the activities and harvests of local fishers is a key method for monitoring coral reefs, 
as it offers insights into ecosystem dynamics and possible pressures. 
This report presents `r tolower(timeframe)` findings from fishery-dependent monitoring of queen conch (_Aliger gigas_) 
at Turneffe Atoll Marine Reserve (TAMR) from `r study_timeframe`. 
Data were recorded by TASA patrolling officers in direct communication with TAMR fishers during their fishing trips. 
Data were collected on key economic species. 
This report aims to summarize and visualize the Fisheries 
Conch Catch dataset from `r study_timeframe` to facilitate data interpretation and strengthen the scientific basis for 
management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Fisheries Catch Project

### Queen Conch

Queen conch (_Aliger gigas_) are large-sized Caribbean gastropods characterized by their distinctive whorl-shaped shell with a set of thick spines at the apex. 
Juvenile queen conch have a pink, pearly interior, and the shell has a recognizable dark periostracum layer. 
Sex is determined by the presence of either a verge or egg groove inside the shell, which can only be observed by removing the conch from its shell. 
Reproduction occurs during warmer months (Appeldoorn and Baker 2013). 

In Belize, the queen conch fishery is the second most important commercial fishery in the country. 
Conch are harvested by free diving depths of 5-60ft in seagrass beds and coral reef habitats. 
The fishery is regulated by the Belize Fisheries Department (BFD). 
National regulations mandate that conch may only be harvested if they exceed a shell length of 7 inches, and only during the open season, which extends from October 1st until June 30th of the following year, or until the national conch quota is reached, whichever occurs first. 
The use of SCUBA for commercial conch harvesting is strictly prohibited. 
Additionally, commercial fishing activities in Belize require a fishing license (Belize Fisheries Department accessed 2025).

According to NOAA, queen conch density needs to be at least 100 conch per hectare to facilitate proper mating (NOAA Fisheries 2025).\


## Methods

This report, including text, figures, and tables, was generated using the 
Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025, version `r params$release`), an R Shiny app which takes 
standardized reef monitoring data, validates the dataset, and creates a shareable output. \


## Results



```{r table1_prep, echo = FALSE, include = FALSE}
# Create df for Table 1 ---------------------------
df_events_overall <- df %>%
  select(`Mission Id`, `Hours Fished`) %>%
  distinct() %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  filter(!is.na(`Hours Fished`), `Hours Fished` > 0) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE)
  ) %>%
  mutate(Month = "Overall")
df_table1_overall <- df %>%
  summarize(
    Vessels = n_distinct(`Name of Vessel`),
    Conch = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  mutate(Month = "Overall") %>%
  left_join(df_events_overall, by = "Month") %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Conch / Hours, 1), NA)
  )
df_events <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  select(Month, `Mission Id`, `Hours Fished`) %>%
  distinct() %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  filter(!is.na(`Hours Fished`), `Hours Fished` > 0) %>%
  group_by(Month) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE)
  )
df_table1 <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Vessels = n_distinct(`Name of Vessel`),
    Conch = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Conch / Hours, 1), NA)
  ) %>%
  bind_rows(df_table1_overall)

# Prepare dynamic variables for Table 1 ---------------------------
total_vessels <- length(unique(df$`Name of Vessel`))
total_conch <- df_table1_overall$Conch
total_events <- df_table1_overall$Missions
total_hours <- df_table1_overall$Hours
max_cpue_num <- max(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
max_cpue_num_month <- df_table1$Month[which.max(df_table1$`CPUE (num/hr)`)]
min_cpue_num <- min(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
min_cpue_num_month <- df_table1$Month[which.min(df_table1$`CPUE (num/hr)`)]
```


Between `r study_start` and `r study_end`, `r total_conch` conch were 
caught and recorded as part of the TAMR 
Fisheries Catch dataset. These conch were caught across `r total_events` missions, 
between `r total_vessels` surveyed vessels, 
totalling `r total_hours` active fishing hours.
The highest CPUE in count 
per hour (num/hr) occurred in `r max_cpue_num_month` 
with `r max_cpue_num`, whereas the lowest was `r min_cpue_num_month` 
in `r min_cpue_num` (Table 1; Figure 1).\

**Table 1.** Summary of fishing activity by month, including number of surveyed vessels, 
missions, and conch caught, along with fishing hours, pounds of catch, 
and catch per unit effort (CPUE). 
CPUE values are calculated only for missions where fishing hours and catch weight were recorded. 
Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\


```{r table1, echo = FALSE}
# show table
as.data.frame(df_table1)
```


<br>


```{r table1_expanded_prep, echo = FALSE, include = FALSE}
# Create df for Table 1 Expanded ---------------------------
df_table1_expanded_rows <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  group_by(`Mission Id`) %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  summarize(
    Month = first(Month),
    Month_sort = first(Month_sort),
    Hours = first(`Hours Fished`),
    Conch = n(),
    .groups = "drop"
  ) %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, Conch / Hours, NA)
  )
df_table1_expanded_num_overall <- df_table1_expanded_rows %>%
  summarize(
    Min = round(min(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (num/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (num/hr)`, na.rm = TRUE), 1)
  ) %>%
  mutate(Month = "Overall", CPUE = "num/hr")
df_table1expanded <- df_table1_expanded_rows %>%
  mutate(CPUE = "num/hr") %>%
  group_by(CPUE, Month, Month_sort) %>%
  summarize(
    Min = round(min(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (num/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (num/hr)`, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  bind_rows(df_table1_expanded_num_overall)

# Prepare dynamic variables for Table 1 Expanded ---------------------------
overall_mean <- df_table1_expanded_num_overall$Mean
overall_median <- df_table1_expanded_num_overall$Median
overall_min <- df_table1_expanded_num_overall$Min
overall_max <- df_table1_expanded_num_overall$Max
overall_sd <- df_table1_expanded_num_overall$SD
max_cpue_mean_num <- max(df_table1expanded$Mean, na.rm = TRUE)
max_cpue_mean_num_month <- df_table1expanded$Month[which.max(df_table1expanded$Mean)]
min_cpue_mean_num <- min(df_table1expanded$Mean, na.rm = TRUE)
min_cpue_mean_num_month <- df_table1expanded$Month[which.min(df_table1expanded$Mean)]
```


The CPUE of conch in count per hour had a 
mean of `r overall_mean`num/hr and median of `r overall_median`num/hr, 
with the highest mean of `r max_cpue_mean_num`num/hr being in `r max_cpue_mean_num_month` 
and the lowest mean of `r min_cpue_mean_num`num/hr being in `r min_cpue_mean_num_month`. 
Minimum CPUE recorded across all missions was `r overall_min`num/hr, maximum was `r overall_max`num/hr, 
and standard deviation was `r overall_sd`num/hr 
(Table 1; Figure 1).\

