---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  release: "Unknown"
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  lamp_1per_year_selection: "None"
  lamp_1per_period_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(moments)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Update Site ID ---------------------------
update_site_id <- function(df) {
  df |> mutate(`Site ID` = ifelse(is.na(`Site ID`) | `Site ID` == "", as.character(`Old Site ID`), as.character(`Site ID`)))
}
df_surveys <- update_site_id(params$datafile$Survey_Data)
df_sites <- update_site_id(params$datafile$Sites)

# Merge sites and surveys data ---------------------------
df_merged <- df_surveys |>
  left_join(df_sites, by = "Site ID") |>
  filter(Section != "NW") |>
  mutate(`Conch Count` = as.integer(`Conch Count`))
if (as.character(params$lamp_1per_year_selection) != "None") {
  df_merged <- df_merged |>
    filter(format(Date.x, "%Y") == params$lamp_1per_year_selection) |>
    filter(format(Date.y, "%Y") == params$lamp_1per_year_selection)
  df_sites <- df_sites |>
    filter(format(Date, "%Y") == params$lamp_1per_year_selection)
  df_surveys <- df_surveys |>
    filter(format(Date, "%Y") == params$lamp_1per_year_selection)
}

# Count number of sites surveyed ---------------------------
number_sites <- length(unique(df_merged$`Site ID`))

# Define study period ---------------------------
study_start <- min(df_merged$Date.x, na.rm = TRUE)
study_end <- max(df_merged$Date.x, na.rm = TRUE)
study_year <- format(study_start, "%Y")

# Define samba conch ---------------------------
samba_definition <- "Samba conch, which may be either a phenotype or subspecies of queen conch, are here defined by being less than 7inches in length with a flared lip of 5mm or more."
```

# Turneffe Atoll Queen Conch LAMP Dataset Report: `r study_year`


Turneffe Atoll Sustainability Association (TASA)


**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\
**App Version:** `r params$release`\

***


Coral reefs are biodiversity hotspots that provide critical habitats for diverse species, support coastal communities, and offer economic opportunities through the harvest of marine invertebrates and fish. 
Monitoring these ecosystems is essential to assess health, function, and changes over time, informing management decisions. 
This report presents findings from monitoring surveys of queen conch (_Aliger gigas_) at Turneffe Atoll Marine Reserve (TAMR) from `r study_year`. 
Data on conch abundance and shell measurements were collected by TASA staff and collaborators as part of the Long-term Atoll Monitoring Program (LAMP), which tracks key taxa at the atoll through standardized surveys. 
This report aims to summarize and visualize the queen conch dataset from `r study_year` to facilitate data interpretation and support management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\


### Long-term Atoll Monitoring Program

The Long-term Atoll Monitoring Program (LAMP) at TAMR, adapted from the original LAMP protocol for Glover’s Reef (Acosta 2004), is designed to monitor key fished taxa at the atoll over space and time, including finfish, queen conch (_Aliger gigas_) and milk conch (_Strombus costatus_), Caribbean spiny lobster (_Panulirus argus_) and spotted spiny lobster (_Panulirus guttatus_), black sea urchins (_Diadema antillarum_), and Caribbean king crab (_Mithrax spinosissumus_). 
The program uses habitat sampling to estimate population sizes and physical characteristics of these taxa, enabling spatial and temporal comparison.

LAMP's primary objectives are to assess the viability of these populations by identifying trends in abundance, reproduction, physical characteristics, and habitat quality where applicable. 
Data collected through LAMP facilitates evaluation of management zone effectiveness. 
The program data also supports recommendations on enforcement, monitoring, education, outreach, fishing quotas, season lengths, size limits, and other government regulations to ensure the long-term sustainability of the fishery, all while preserving the natural beauty and ecological value of the atoll.

The queen conch dataset is a key component of the overall LAMP data, notable for its extensive historical record and the breadth of information collected. 
Queen conch is a commercially significant species within the TAMR. 
As outlined in the Turneffe Atoll Management Plan, regular population monitoring is required to support sustainable management and conservation efforts. 
In 2013, Zale et al. developed a methodology based on Acosta’s LAMP protocol (2004) to monitor queen conch at TAMR, which was later reused in 2016 and 2017 to assess changes in the stock over time and improve the monitoring methodology (Zale et al. 2016, Zale et al. 2018). 
From 2023 onward, these datasets are integrated into the broader LAMP program data, which includes crab, lobster, urchins, and finfish, to enhance the program. 
Data from monitoring surveys of queen conch at TAMR include conch abundance, depth of individuals, and shell measurements (lip thickness and shell length).\


### Queen Conch

Queen conch (_Aliger gigas_) are large-sized Caribbean gastropods characterized by their distinctive whorl-shaped shell with a set of thick spines at the apex. 
Juvenile queen conch have a pink, pearly interior, and the shell has a recognizable dark periostracum layer. 
Sex is determined by the presence of either a verge or egg groove inside the shell, which can only be observed by removing the conch from its shell. 
Reproduction occurs during warmer months (Appeldoorn and Baker 2013). 
`r samba_definition`

In Belize, the queen conch fishery is the second most important commercial fishery in the country. 
Conch are harvested by free diving depths of 5-60ft in seagrass beds and coral reef habitats. 
The fishery is regulated by the Belize Fisheries Department (BFD). 
National regulations mandate that conch may only be harvested if they exceed a shell length of 7 inches, and only during the open season, which extends from October 1st until June 30th of the following year, or until the national conch quota is reached, whichever occurs first. 
The use of SCUBA for commercial conch harvesting is strictly prohibited. 
Additionally, commercial fishing activities in Belize require a fishing license (Belize Fisheries Department accessed 2025).

According to NOAA, queen conch density needs to be at least 100 conch per hectare to facilitate proper mating (NOAA Fisheries 2025).\


## Methods

Survey sites were selected as representative nearshore habitat types (as defined by Mumby and Harborne, 1999) known to be inhabited by Queen Conch based on prior assessments (Zale et al., 2016). 
Each site is assigned a unique ID and decimal coordinates to facilitate resampling in future years. 
When a site is re-sampled, habitat classification is reevaluated for the corresponding sampling period. 
Additionally, each site is categorized within a specific section of TAMR, either North, Northeast, Southeast, or Southwest (Figure 1).

The survey period for this report’s dataset extended from `r study_start` to `r study_end` at TAMR. 
Sampling for queen conch LAMP is ideally done during summer before harvest. `r number_sites` sites were sampled in total. 
At each site, four 50m by 4m belt transects were surveyed by data collectors either diving, snorkeling, or wading depending on the depth. 
Transects were placed parallel to each other 10m apart and parallel to wind direction where possible. 
Each transect line was established using one white 3/16” non-stretchable sinking Dacron rope, which was pulled taut and anchored at either end with buoy lines on each anchor extended to the surface for retrieval. 
The belt transect included 2m of space on either side of the line.

During sampling, two data collectors swam together down either side of each transect approximately 1.5m above the substrate. 
In sites shallower than 1.5m in depth, data collectors instead snorkeled, and in sites less than 0.25m in depth, data collectors waded. 
The data collectors counted all living queen conch within 2m of the line on their side of the belt. 
Queen conch close to 2m from the line were also counted, and their distance from the line recorded. 
Shell lengths and lip thickness of queen conch were measured during sampling using calipers, as well as depth of capture to the nearest foot using a dive computer. 
Lip thickness was measured only in lipped individuals, and was measured as 4cm inward from the edge of the middle third of the lip, where the lip is relatively straight instead of wavy. 
After being measured, all queen conch were returned upright to their original position of capture.

This report, including text, figures, and tables, was generated using the Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025, version `r params$release`), an R Shiny app which takes standardized reef monitoring data, validates the dataset, and creates a shareable output. \


```{r figure1, echo = FALSE, fig.width = 7, fig.height = 7}
# Create map for Figure 1 ---------------------------
df_map <- df_merged |>
  select(`Site ID`, Latitude, Longitude) |>
  distinct()
if (length(unique(df_merged$`Site ID`)) < 50) {
  generate_map(df_map, "Latitude", "Longitude", "Site ID")
} else {
  generate_map_nolabels(df_map, "Latitude", "Longitude", "Site ID")
}
```


*Figure 1.* Map of Turneffe Atoll Marine Reserve zones and ecosystems. Shapefiles for zones belong to the Belize Fisheries Department (Belize Geomatics 2022). 
Shapefiles for the base map and ecosystems come from the publicly available Spatial Data Warehouse for Belize (Meerman 2013, Meerman 2017).
Sample sites surveyed in `r study_year` are labelled on the map.\


## Results

```{r table1_prep, echo = FALSE, include = FALSE}
# Create df for Table 1 ---------------------------
df_table1 <- df_merged %>%
  group_by(Section) %>%
  summarise(
    `Number Sites` = length(unique(`Site ID`)),
    `Queen Conch Count` = sum(`Conch Count`, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    `Section Hectares` = `Number Sites` * 0.08,
    `Conch / Hectare` = round(`Queen Conch Count` / `Section Hectares`, 0)
  ) %>%
  bind_rows(
    df_merged %>%
      summarise(
        Section = "Total",
        `Number Sites` = length(unique(`Site ID`)),
        `Queen Conch Count` = sum(`Conch Count`, na.rm = TRUE)
      ) %>%
      mutate(
        `Section Hectares` = `Number Sites` * 0.08,
        `Conch / Hectare` = round(`Queen Conch Count` / `Section Hectares`, 0)
      )
  )

# Prepare dynamic variables for Table 1 ---------------------------
total_conch <- filter(df_table1, Section == "Total")$`Queen Conch Count`
total_sites <- filter(df_table1, Section == "Total")$`Number Sites`
NO_sites <- filter(df_table1, Section == "NO")$`Number Sites`
NE_sites <- filter(df_table1, Section == "NE")$`Number Sites`
SW_sites <- filter(df_table1, Section == "SW")$`Number Sites`
SE_sites <- filter(df_table1, Section == "SE")$`Number Sites`
```


In total, out of `r total_sites` sampled sites, `r total_conch` queen conch were counted. 
`r NE_sites` sites were sampled in the Northeast, `r NO_sites` in the North, `r SE_sites` in the Southeast, and `r SW_sites` in the Southwest (Table 1).\


**Table 1.**  Table of summary data for sampled sections of Turneffe Atoll Marine Reserve for data from `r study_year`. 
NO is north, NE is northeast, SE is southeast, and SW is southwest.\

```{r table1, echo = FALSE}
df_table1
```

<br>

```{r table2_prep, echo = FALSE, include = FALSE}
# Create df for Table 2 ---------------------------
df_sites_summary <- df_merged %>%
  group_by(Section, `Site ID`) %>%
  summarise(`Site Conch Count` = sum(`Conch Count`), .groups = "drop")
df_table2 <- df_sites_summary %>%
  group_by(Section) %>%
  summarise(
    Mean = round(mean(`Site Conch Count`), 1),
    Min = round(min(`Site Conch Count`), 1),
    Q1 = round(quantile(`Site Conch Count`, 0.25), 1),
    Median = round(median(`Site Conch Count`), 1),
    Q3 = round(quantile(`Site Conch Count`, 0.75), 1),
    Max = round(max(`Site Conch Count`), 1),
    Var = round(var(`Site Conch Count`), 1),
    SD = round(sd(`Site Conch Count`), 1),
    .groups = "drop"
  ) %>%
  bind_rows(
    summarise(
      df_sites_summary,
      Section = "Total",
      Mean = round(mean(`Site Conch Count`), 1),
      Min = round(min(`Site Conch Count`), 1),
      Q1 = round(quantile(`Site Conch Count`, 0.25), 1),
      Median = round(median(`Site Conch Count`), 1),
      Q3 = round(quantile(`Site Conch Count`, 0.75), 1),
      Max = round(max(`Site Conch Count`), 1),
      Var = round(var(`Site Conch Count`), 1),
      SD = round(sd(`Site Conch Count`), 1)
    ) %>%
      distinct(Section, Mean, Min, Q1, Median, Q3, Max, Var, SD)
  )

# Prepare dynamic variables for Table 2 ---------------------------
overall_skewness <- round(skewness(df_sites_summary$`Site Conch Count`), 2)
overall_kurtosis <- round(kurtosis(df_sites_summary$`Site Conch Count`), 2)
skew_direction <- ifelse(overall_skewness > 0, "positive", "negative")
skew_intensity <- case_when(abs(overall_skewness) > 3 ~ "highly", abs(overall_skewness) > 1 ~ "moderately", abs(overall_skewness) <= 1 ~ "slightly")
skew_means_interpretation <- case_when(
  skew_direction == "positive" & skew_intensity == "highly" ~ "much higher",
  skew_direction == "negative" & skew_intensity == "highly" ~ "much lower",
  skew_direction == "positive" ~ "higher",
  skew_direction == "negative" ~ "lower"
)
skew_presence_interpretation <- case_when(
  skew_direction == "positive" & skew_intensity == "highly" ~ "absent or found in low numbers",
  skew_direction == "positive" ~ "found in low numbers",
  skew_direction == "negative" ~ "found in high numbers"
)
kurtosis_classification <- case_when(
  overall_kurtosis > 2.5 & overall_kurtosis < 3.5 ~ "normal distribution",
  overall_kurtosis > 10 ~ "highly leptokurtic distribution, meaning there are far more extreme outliers than a normal distribution",
  overall_kurtosis >= 3.5 ~ "leptokurtic distribution, meaning there are more extreme outliers than a normal distribution",
  overall_kurtosis <= 2.5 ~ "platykurtic distribution, meaning there are fewer outliers than a normal distribution"
)
kurtosis_presence_interpretation <- case_when(
  overall_kurtosis > 2.5 & overall_kurtosis < 3.5 ~ "Due to the normal distribution, most conch counts are expected to follow that pattern with some outliers",
  overall_kurtosis > 10 ~ "However, due to the highly leptokurtic distribution, extreme outliers in counts are expected, potentially many times the median in value",
  overall_kurtosis >= 3.5 ~ "However, due to the leptokurtic distribution, outliers in counts are expected, potentially many times the median in value",
  overall_kurtosis <= 2.5 ~ "Due to the platykurtic distribution, most conch counts are expected to follow that pattern with few outliers"
)
maximum_conch <- max(df_sites_summary$`Site Conch Count`, na.rm = TRUE)
maximum_conch_site <- df_sites_summary$`Site ID`[which.max(df_sites_summary$`Site Conch Count`)]
highest_median_conch <- max(df_table2$Median, na.rm = TRUE)
highest_median_conch_section <- df_table2 %>%
  filter(Median == highest_median_conch) %>%
  pull(Section) %>%
  unique()
highest_median_conch_section_plural <- ifelse(length(highest_median_conch_section) > 1, "s", "")
highest_median_conch_section_str <- knitr::combine_words(highest_median_conch_section)
overall_median_conch <- filter(df_table2, Section == "Total")$Median
```


The distribution of Queen Conch counts has a skew of `r overall_skewness` and a kurtosis of `r overall_kurtosis`.
This indicates that Queen Conch counts follow a `r kurtosis_classification`, `r skew_intensity` skewed towards the `r skew_direction` direction.
This leads to the means of Queen Conch counts to be `r skew_means_interpretation` than the median.
This suggests that Queen Conch are typically `r skew_presence_interpretation` at Turneffe sites. `r kurtosis_presence_interpretation` (Table 2; Figure 2). 
The maximum queen conch count of `r maximum_conch` was found at site `r maximum_conch_site`. 
The highest median queen conch count of `r highest_median_conch` per site was found in the `r highest_median_conch_section_str` section`r highest_median_conch_section_plural`, compared to the overall median queen conch count of `r overall_median_conch` (Table 2). \


```{r figure2_prep, echo = FALSE, include = FALSE}
# Create summary dfs for Figure 2 ---------------------------
summary_transect <- df_merged %>%
  select(Date.x, `Site ID`, Transect, Section, `Conch Count`, `Shell Length (in)`, `Lip Thickness (mm)`) %>%
  group_by(Date.x, `Site ID`, Transect, Section) %>%
  summarize(Conch_Count_Transect = sum(`Conch Count`)) %>%
  ungroup()
summary_site <- summary_transect %>%
  group_by(`Site ID`, Section) %>%
  summarize(
    Num_Transects = n(),
    Conch_Count_Site = sum(Conch_Count_Transect),
    Mean_Conch_Per_Transect = mean(Conch_Count_Transect),
    Median_Conch_Per_Transect = median(Conch_Count_Transect),
    Min_Conch_Per_Transect = min(Conch_Count_Transect),
    Max_Conch_Per_Transect = max(Conch_Count_Transect)
  )
mean_per_transect <- summary_transect %>%
  group_by(Section) %>%
  summarize(Mean_Conch_Per_Transect = mean(Conch_Count_Transect))
mean_per_site <- summary_site %>%
  group_by(Section) %>%
  summarize(Mean_Conch_Per_Site = mean(Conch_Count_Site))

# Create panels for FIgure 2 ---------------------------
A <- ggplot(summary_transect, aes(x = Conch_Count_Transect)) +
  geom_histogram(
    binwidth = 1, color = "black",
    fill = "white", boundary = 0
  ) +
  geom_boxplot(aes(y = 175), width = 10, size = 1) +
  geom_vline(
    xintercept = mean(summary_transect$Conch_Count_Transect),
    size = 1, linetype = "dashed"
  ) +
  theme_custom +
  xlab("Conch per Transect") +
  ylab("Number of Transects") +
  labs(title = "A") +
  scale_x_continuous(
    breaks = c(0, 20, 40, 60, 80, 100),
    labels = c(0, 20, 40, 60, 80, 100)
  ) +
  coord_cartesian(xlim = c(1, 110))
B <- ggplot(summary_site, aes(x = Conch_Count_Site)) +
  geom_histogram(
    binwidth = 5, color = "black",
    fill = "white", boundary = 0
  ) +
  geom_boxplot(aes(y = 45), width = 3.5, size = 1) +
  geom_vline(
    xintercept = mean(summary_site$Conch_Count_Site),
    size = 1, linetype = "dashed"
  ) +
  theme_custom +
  xlab("Conch per Site") +
  ylab("Number of Sites") +
  labs(title = "B") +
  scale_x_continuous(
    breaks = c(0, 50, 100, 150, 200, 250),
    labels = c(0, 50, 100, 150, 200, 250)
  ) +
  coord_cartesian(xlim = c(1, 275))
C <- ggplot(
  summary_transect,
  aes(x = Section, y = Conch_Count_Transect, fill = Section)
) +
  geom_boxplot(fill = "white") +
  geom_point(
    data = mean_per_transect,
    aes(x = Section, y = Mean_Conch_Per_Transect),
    size = 5, shape = 2, stroke = 1
  ) +
  theme_custom +
  ylab("Conch per Transect") +
  guides(fill = "none") +
  labs(title = "C")
D <- ggplot(
  summary_site,
  aes(x = Section, y = Conch_Count_Site, fill = Section)
) +
  geom_boxplot(fill = "white") +
  geom_point(
    data = mean_per_site,
    aes(x = Section, y = Mean_Conch_Per_Site),
    size = 5, shape = 2, stroke = 1
  ) +
  theme_custom +
  ylab("Conch per Site") +
  guides(fill = "none") +
  labs(title = "D")

# Assemble Figure 2 ---------------------------
figure2_bottom <- ggarrange(C, D, ncol = 2, nrow = 1)
figure2 <- ggarrange(A, B, figure2_bottom, ncol = 1, nrow = 3)
```


```{r figure2, fig.width=6, fig.height=7.5, fig.align = 'left', echo = FALSE}
figure2
```


*Figure 2.* Distribution of queen conch counts for data from `r study_year` sampling effort at Turneffe Atoll Marine Reserve. 
Histogram and boxplot of conch counts per transect (bin size = 1) with mean shown by dotted line (A), 
histogram and boxplot of conch counts per site (bin size = 5) with mean shown by dotted line (B), 
boxplots of conch counts per transect divided by section with means indicated with triangles (C), 
and boxplots of conch counts per site divided by section with means indicated with triangles (D). 
For all boxplots, the box encompasses 50% of the data, and the whiskers span the calculated minima to maxima. 
The bar inside the box represents the median.\



**Table 2.**  Table of summary statistics for conch counts per site in sections of North, Northeast, Southeast, and Southwest Turneffe Atoll Marine Reserve for data from `r study_year`. 
NO is north, NE is northeast, SE is southeast, and SW is southwest. Summary statistics include mean, minimum, first quartile, median, third quartile, maximum, variance, and standard deviation.\


```{r table2, echo = FALSE}
df_table2
```

<br>


```{r table3_prep, echo = FALSE, include = FALSE}
# Create df for Table 3 ---------------------------
df_table3_percent <- df_merged %>%
  group_by(Section) %>%
  summarize(
    Type = "(%)",
    `Legal (≥ 17.8 cm)` = round(100 * sum(`Conch Count`[`Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Sublegal (< 17.8 cm)` = round(100 * sum(`Conch Count`[`Shell Length (in)` < 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Lipped` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` > 0], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Possibly Mature (≥ 4 mm lip)` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 4], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Probably Mature (≥ 15 mm lip)` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 15], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Legal, Lipped` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Legal, Possibly Mature` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 4 & `Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Legal, Probably Mature` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 15 & `Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Sublegal, Lipped` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` < 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Samba (<17.78cm, 5mm+ lip)` = round(100 * sum(`Conch Count`[
      `Lip Thickness (mm)` >= 5 &
        `Shell Length (in)` <= 7
    ], na.rm = TRUE) / sum(`Conch Count`), 0)
  ) %>%
  pivot_longer(
    cols = -c(Section, Type),
    names_to = "Attribute",
    values_to = "Value"
  )
df_table3_counts <- df_merged %>%
  group_by(Section) %>%
  summarize(
    Type = "(#)",
    `Legal (≥ 17.8 cm)` = round(sum(`Conch Count`[`Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Sublegal (< 17.8 cm)` = round(sum(`Conch Count`[`Shell Length (in)` < 7], na.rm = TRUE), 1),
    `Lipped` = round(sum(`Conch Count`[`Lip Thickness (mm)` > 0], na.rm = TRUE), 1),
    `Possibly Mature (≥ 4 mm lip)` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 4], na.rm = TRUE), 1),
    `Probably Mature (≥ 15 mm lip)` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 15], na.rm = TRUE), 1),
    `Legal, Lipped` = round(sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Legal, Possibly Mature` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 4 & `Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Legal, Probably Mature` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 15 & `Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Sublegal, Lipped` = round(sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` < 7], na.rm = TRUE), 1),
    `Samba (<17.78cm, 5mm+ lip)` = round(sum(`Conch Count`[
      `Lip Thickness (mm)` >= 5 &
        `Shell Length (in)` <= 7
    ], na.rm = TRUE), 1)
  ) %>%
  pivot_longer(
    cols = -c(Section, Type),
    names_to = "Attribute",
    values_to = "Value"
  )
df_table3_total_percent <- df_merged %>%
  summarize(
    Section = "Overall",
    Type = "(%)",
    `Legal (≥ 17.8 cm)` = round(100 * sum(`Conch Count`[`Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Sublegal (< 17.8 cm)` = round(100 * sum(`Conch Count`[`Shell Length (in)` < 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Lipped` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` > 0], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Possibly Mature (≥ 4 mm lip)` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 4], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Probably Mature (≥ 15 mm lip)` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 15], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Legal, Lipped` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Legal, Possibly Mature` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 4 & `Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Legal, Probably Mature` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` >= 15 & `Shell Length (in)` >= 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Sublegal, Lipped` = round(100 * sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` < 7], na.rm = TRUE) / sum(`Conch Count`), 0),
    `Samba (<17.78cm, 5mm+ lip)` = round(100 * sum(`Conch Count`[
      `Lip Thickness (mm)` >= 5 &
        `Shell Length (in)` <= 7
    ], na.rm = TRUE) / sum(`Conch Count`), 0)
  ) %>%
  pivot_longer(
    cols = -c(Section, Type),
    names_to = "Attribute",
    values_to = "Value"
  )
df_table3_total_counts <- df_merged %>%
  summarize(
    Section = "Overall",
    Type = "(#)",
    `Legal (≥ 17.8 cm)` = round(sum(`Conch Count`[`Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Sublegal (< 17.8 cm)` = round(sum(`Conch Count`[`Shell Length (in)` < 7], na.rm = TRUE), 1),
    `Lipped` = round(sum(`Conch Count`[`Lip Thickness (mm)` > 0], na.rm = TRUE), 1),
    `Possibly Mature (≥ 4 mm lip)` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 4], na.rm = TRUE), 1),
    `Probably Mature (≥ 15 mm lip)` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 15], na.rm = TRUE), 1),
    `Legal, Lipped` = round(sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Legal, Possibly Mature` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 4 & `Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Legal, Probably Mature` = round(sum(`Conch Count`[`Lip Thickness (mm)` >= 15 & `Shell Length (in)` >= 7], na.rm = TRUE), 1),
    `Sublegal, Lipped` = round(sum(`Conch Count`[`Lip Thickness (mm)` > 0 & `Shell Length (in)` < 7], na.rm = TRUE), 1),
    `Samba (<17.78cm, 5mm+ lip)` = round(sum(`Conch Count`[
      `Lip Thickness (mm)` >= 5 &
        `Shell Length (in)` <= 7
    ], na.rm = TRUE), 1)
  ) %>%
  pivot_longer(
    cols = -c(Section, Type),
    names_to = "Attribute",
    values_to = "Value"
  )
df_table3_bound <- bind_rows(df_table3_percent, df_table3_counts, df_table3_total_percent, df_table3_total_counts)
df_table3 <- df_table3_bound %>%
  pivot_wider(names_from = Type, values_from = Value) %>%
  mutate(Combined = paste0(`(%)`, "%, (", `(#)`, ")")) %>%
  select(Section, Attribute, Combined) %>%
  pivot_wider(names_from = Section, values_from = Combined)

# Prepare dynamic variables for Table 3 ---------------------------
maximum_legal <- max(filter(df_table3_percent, Attribute == "Legal (≥ 17.8 cm)")$Value, na.rm = TRUE)
maximum_legal_section <- filter(df_table3_percent, Attribute == "Legal (≥ 17.8 cm)")$Section[which.max(filter(df_table3_percent, Attribute == "Legal (≥ 17.8 cm)")$Value)]
legal_overall <- filter(df_table3_total_percent, Attribute == "Legal (≥ 17.8 cm)")$Value
maximum_lipped <- max(filter(df_table3_percent, Attribute == "Lipped")$Value, na.rm = TRUE)
maximum_lipped_section <- filter(df_table3_percent, Attribute == "Lipped")$Section[which.max(filter(df_table3_percent, Attribute == "Lipped")$Value)]
lipped_overall <- filter(df_table3_total_percent, Attribute == "Lipped")$Value
maximum_legal_lipped <- max(filter(df_table3_percent, Attribute == "Legal, Lipped")$Value, na.rm = TRUE)
maximum_legal_lipped_section <- filter(df_table3_percent, Attribute == "Legal, Lipped")$Section[which.max(filter(df_table3_percent, Attribute == "Legal, Lipped")$Value)]
legal_lipped_overall <- filter(df_table3_total_percent, Attribute == "Legal, Lipped")$Value
```


The highest proportion of legal (≥ 17.8 cm shell length) queen conch was found in the `r maximum_legal_section` section with `r maximum_legal`%, compared to the overall proportion of `r legal_overall`%. 
The highest proportion of lipped queen conch was found in the `r maximum_lipped_section` with `r maximum_lipped`%, compared to the overall proportion of `r lipped_overall`%. 
The highest proportion of legal and lipped queen conch was found in the `r maximum_legal_lipped_section` with `r maximum_legal_lipped`%, compared to the overall proportion of `r legal_lipped_overall`% (Table 3; Figure 3). \


```{r figure3_prep, echo = FALSE, include = FALSE}
# Create df for Figure 3 ---------------------------
df_fig3 <- df_merged %>%
  mutate(
    `Shell Length (cm)` = `Shell Length (in)` * 2.54,
    Lipped = factor(
      case_when(
        is.na(`Lip Thickness (mm)`) ~ "Not Lipped",
        `Lip Thickness (mm)` > 0 ~ "Lipped",
        TRUE ~ "Not Lipped"
      ),
      levels = c("Not Lipped", "Lipped")
    )
  )

# Create panels for Figure 3 ---------------------------
A <- ggplot(filter(df_fig3, Section == "NE"), aes(x = `Shell Length (cm)`, fill = Lipped)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  labs(title = "A", subtitle = "Northeast") +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5))
B <- ggplot(filter(df_fig3, Section == "NO"), aes(x = `Shell Length (cm)`, fill = Lipped)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  labs(title = "B", subtitle = "North") +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5))
C <- ggplot(filter(df_fig3, Section == "SE"), aes(x = `Shell Length (cm)`, fill = Lipped)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  labs(title = "C", subtitle = "Southeast") +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5))
D <- ggplot(filter(df_fig3, Section == "SW"), aes(x = `Shell Length (cm)`, fill = Lipped)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  labs(title = "D", subtitle = "Southwest") +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5))
E <- ggplot(df_fig3, aes(x = `Shell Length (cm)`, fill = Lipped)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  labs(title = "E", subtitle = "Overall") +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5))

# Assemble Figure 3 ---------------------------
figure3 <- ggarrange(A, B, C, D, E, nrow = 3, ncol = 2)
```


```{r figure3, fig.width=6, fig.height=8, fig.align = 'left', echo = FALSE}
figure3
```


*Figure 3.* Length-frequency distribution of lipped and non-lipped queen conch shell length (cm)  for data from `r study_year` in the northeast (A),
north (B),southeast (C), southwest (D), and overall (E) at Turneffe Atoll Marine Reserve. 
The dotted vertical lines indicate the legal shell length (17.8 cm) for harvest.\

<br>


**Table 3.**  Table of queen conch survey attributes for sampled sections of North, Northeast, Southeast, and Southwest Turneffe Atoll Marine Reserve as well as overall for data from `r study_year`.
 NO is north, NE is northeast, SE is southeast, and SW is southwest. 
 Values for each attribute are presented as a percentage of total conch from the section, as well as count.\


```{r table3, echo = FALSE}
df_table3
```


<br>


```{r table4_prep, echo = FALSE, include = FALSE}
# Create df for Table 4 ---------------------------
df_table4_lengths <- df_merged %>%
  group_by(Section) %>%
  summarize(
    `Mean Shell Length` = round(mean(`Shell Length (in)`, na.rm = TRUE) * 2.54, 1),
    `Mean Shell Length (Lipped)` = round(mean(`Shell Length (in)`[`Lip Thickness (mm)` > 0], na.rm = TRUE) * 2.54, 1),
    `Mean Shell Length (non-Lipped)` = round(mean(`Shell Length (in)`[`Lip Thickness (mm)` == 0 | is.na(`Lip Thickness (mm)`)], na.rm = TRUE) * 2.54, 1),
    `Median Shell Length` = round(median(`Shell Length (in)`, na.rm = TRUE) * 2.54, 1)
  ) %>%
  select(Section, `Mean Shell Length`, `Mean Shell Length (Lipped)`, `Mean Shell Length (non-Lipped)`, `Median Shell Length`)

df_table4_total_lengths <- df_merged %>%
  mutate(
    Section = "Overall",
    `Mean Shell Length` = round(mean(`Shell Length (in)`, na.rm = TRUE) * 2.54, 1),
    `Mean Shell Length (Lipped)` = round(mean(`Shell Length (in)`[`Lip Thickness (mm)` > 0], na.rm = TRUE) * 2.54, 1),
    `Mean Shell Length (non-Lipped)` = round(mean(`Shell Length (in)`[`Lip Thickness (mm)` == 0 | is.na(`Lip Thickness (mm)`)], na.rm = TRUE) * 2.54, 1),
    `Median Shell Length` = round(median(`Shell Length (in)`, na.rm = TRUE) * 2.54, 1)
  ) %>%
  select(Section, `Mean Shell Length`, `Mean Shell Length (Lipped)`, `Mean Shell Length (non-Lipped)`, `Median Shell Length`)

df_table4 <- bind_rows(df_table4_lengths, df_table4_total_lengths) %>%
  distinct()

# Prepare dynamic variables for Table 4 ---------------------------
maximum_length <- max(df_table4_lengths$`Mean Shell Length`, na.rm = TRUE)
maximum_length_section <- max(df_table4_lengths$Section[which.max(df_table4_lengths$`Mean Shell Length`)], na.rm = TRUE)
maximum_length_overall <- max(filter(df_table4_total_lengths, Section == "Overall")$`Mean Shell Length`, na.rm = TRUE)
```

The highest mean queen conch shell length was found in the `r maximum_length_section` section, being `r maximum_length`cm. 
This is compared to the overall mean length of `r maximum_length_overall` (Table 4).



**Table 4.**  Table of queen conch shell lengths (cm) for sampled sections of North, Northeast, Southeast, and Southwest Turneffe Atoll Marine Reserve as well as overall for data from `r study_year`.
 NO is north, NE is northeast, SE is southeast, and SW is southwest. \


```{r table4, fig.width=6, fig.height=8, fig.align = 'left', echo = FALSE}
df_table4
```


<br>


## Conclusion

By summarizing and visualizing the queen conch LAMP survey data from `r study_year`, researchers, policy-makers, and other stakeholders are better able to interpret and use the dataset. 
This report supports the aims of the Long-term Atoll Monitoring Program by offering a more understandable view of the data itself, supporting future efforts to create recommendations on enforcement, monitoring, education, outreach, and regulations to ensure the long-term sustainability of the TAMR fishery while safeguarding the ecological value of the atoll.\


## Acknowledgments



```{r collectors_prep, echo = FALSE, include = FALSE}
# Create dynamic variable for acknowledgements ---------------------------
collectors <- df_merged$`Recorder(s)` %>%
  as.character() %>%
  na.omit() %>%
  strsplit(", ") %>%
  unlist()
if (length(collectors) == 0) {
  collectors_string <- ""
} else {
  collectors <- unique(collectors)
  collectors_string <- ifelse(length(collectors) > 1,
    paste0(": ", knitr::combine_words(setdiff(collectors, "Missing"))),
    ""
  )
}
```


Thank you to the data collectors who contributed to this project`r collectors_string`.\


## Literature Cited

Acosta, C. A. (2004). Glover’s Reef long-term atoll monitoring program (LAMP). Glover’s Reef Marine Research Station, Wildlife Conservation Society, Belize.

Appeldoorn, R. S., & Baker, N. (2013). A Literature Review of the Queen Conch, Strombus gigas [Technical Report]. University of Puerto Rico.

Belize Fisheries Department. (n.d.). Capture Fisheries Unit. https://fisheries.gov.bz/units/cfu/

Belize Geomatics. (2022). Turneffe Atoll Areas, Belize Replenishment Zones Project [Dataset].

Egan, B. D. (1985). Aspects of the reproductive biology of Strombus gigas. [Masters Thesis]. University of British Columbia, Vancouver.

Fedler, A. J. (2011). The Economic Value of Turneffe Atoll (Full Report). Turneffe Atoll Trust. http://www.turneffeatollmarinereserve.org/app/webroot/userfiles/214/File/Turneffe%20Atoll%20Valuation%20Study%20PDF%20-%20Final%20Report.pdf

Foley, J. R., & Takahashi, M. (2017). Shell Lip Thickness Is the Most Reliable Proxy to Sexual Maturity in Queen Conch (Lobatus gigas) of Port Honduras Marine Reserve, Belize; Informing Management to Reduce the Risk of Growth Overfishing. Frontiers in Marine Science, 4, 179. https://doi.org/10.3389/fmars.2017.00179

Meerman, J. C. (2013). Belize Basemap [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Meerman, J. C. (2017). Belize_Ecosystems_2017 [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Mumby, P. J., & Harborne, A. R. (1999). Classification scheme for marine habitats of Belize, 5th draft. UNDP/GEF Belize Coastal Zone Management Project.

NOAA Fisheries. (2025). Queen Conch: Science. https://www.fisheries.noaa.gov/species/queen-conch/science

Stoddart, D. R. (1962). Three Caribbean atolls: Turneffe Islands, Lighthouse Reef and Glover’s Reef,  British Honduras. Atoll Research Bulletin, 87, 1–151.

Stoner, A. W., Mueller, K. W., Brown-Peterson, N. J., Davis, M. H., & Booker, C. J. (2012). Maturation and age in queen conch (Strombus gigas): Urgent need for changes in harvest criteria. Fisheries Research, 131–133, 76–84. https://doi.org/10.1016/j.fishres.2012.07.017

Turneffe Atoll Sustainability Association. (2025). Turneffe Reef-Monitoring Data Reporting Tool (Version 1.0) [Computer software].

Zale, A. V., Bramblett, R. G., & Anderson, A. R. (2016). Baseline population abundance estimate and development of a monitoring plan for Queen Conch at Turneffe Atoll Marine Reserve. Submitted to Turneffe Atoll Marine Reserve.

