---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  release: "Unknown"
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  fisheries_1per_timeframe: "Annual"
  fisheries_1per_year_selection: "None"
  fisheries_1per_period_selection: "None"
  fisheries_1per_lobster_season_selection: "None" 
  fisheries_1per_conch_season_selection: "None" 
  fisheries_1per_finfish_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Determine year or season  ---------------------------
timeframe <- params$fisheries_1per_timeframe

# Filter to year or season  ---------------------------
df <- params$datafile[[1]]
df$`Date of Encounter` <- as.Date(df$`Date of Encounter`)
if (timeframe == "Annual" && as.character(params$fisheries_1per_year_selection) != "None") {
  df <- df |>
    filter(format(`Date of Encounter`, "%Y") == params$fisheries_1per_year_selection)
} else if (timeframe == "Seasonal" && as.character(params$fisheries_1per_conch_season_selection) != "None") {
  season_years <- unlist(strsplit(params$fisheries_1per_conch_season_selection, "-"))
  start_date <- as.Date(paste0(season_years[1], "-10-01"))
  end_date <- as.Date(paste0(season_years[2], "-06-28"))
  df <- df |>
    filter(`Date of Encounter` >= start_date, `Date of Encounter` <= end_date)
}

# Add month column  ---------------------------
df$Month <- format(df$`Date of Encounter`, "%b %Y")

# Define study timeframe ---------------------------
study_start <- min(df$`Date of Encounter`, na.rm = TRUE)
study_end <- max(df$`Date of Encounter`, na.rm = TRUE)
if (timeframe == "Annual") {
  study_timeframe <- unique(format(df$`Date of Encounter`, "%Y"))[1]
} else if (timeframe == "Seasonal") {
  study_timeframe <- paste0(format(study_start, "%Y"), "-", format(study_end, "%Y"))
}
```

# Turneffe Atoll Fisheries Catch Conch `r timeframe` Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)

***

Coral reef ecosystems, like those of Turneffe Atoll Marine Reserve (TAMR), provide key ecological and economic value to coastal families. 
With proper management and monitoring, 
these ecosystems can continue to support people sustainably in the long-term. 
Directly tracking the activities and harvests of local fishers is a key method for monitoring coral reefs, 
as it offers insights into ecosystem dynamics and possible pressures. 
This report presents `r tolower(timeframe)` findings from fishery-dependent monitoring of queen conch (_Aliger gigas_) 
at Turneffe Atoll Marine Reserve (TAMR) from `r study_timeframe`. 
Data were recorded by TASA patrolling officers in direct communication with TAMR fishers during their fishing trips. 
Data were collected on key economic species. 
This report aims to summarize and visualize the Fisheries 
Conch Catch dataset from `r study_timeframe` to facilitate data interpretation and strengthen the scientific basis for 
management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Fisheries Catch Project

### Queen Conch

Queen conch (_Aliger gigas_) are large-sized Caribbean gastropods characterized by their distinctive whorl-shaped shell with a set of thick spines at the apex. 
Juvenile queen conch have a pink, pearly interior, and the shell has a recognizable dark periostracum layer. 
Sex is determined by the presence of either a verge or egg groove inside the shell, which can only be observed by removing the conch from its shell. 
Reproduction occurs during warmer months (Appeldoorn and Baker 2013). 

In Belize, the queen conch fishery is the second most important commercial fishery in the country. 
Conch are harvested by free diving depths of 5-60ft in seagrass beds and coral reef habitats. 
The fishery is regulated by the Belize Fisheries Department (BFD). 
National regulations mandate that conch may only be harvested if they exceed a shell length of 7 inches, and only during the open season, which extends from October 1st until June 30th of the following year, or until the national conch quota is reached, whichever occurs first. 
The use of SCUBA for commercial conch harvesting is strictly prohibited. 
Additionally, commercial fishing activities in Belize require a fishing license (Belize Fisheries Department accessed 2025).

According to NOAA, queen conch density needs to be at least 100 conch per hectare to facilitate proper mating (NOAA Fisheries 2025).\


## Methods

This report, including text, figures, and tables, was generated using the 
Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025, version `r params$release`), an R Shiny app which takes 
standardized reef monitoring data, validates the dataset, and creates a shareable output. \


## Results



```{r table1_prep, echo = FALSE, include = FALSE}
# Create df for Table 1 ---------------------------
df_events_overall <- df %>%
  select(`Mission Id`, `Hours Fished`) %>%
  distinct() %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  filter(!is.na(`Hours Fished`), `Hours Fished` > 0) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE)
  ) %>%
  mutate(Month = "Overall")
df_table1_overall <- df %>%
  summarize(
    Vessels = n_distinct(`Name of Vessel`),
    Conch = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  mutate(Month = "Overall") %>%
  left_join(df_events_overall, by = "Month") %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Conch / Hours, 1), NA)
  )
df_events <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  select(Month, `Mission Id`, `Hours Fished`) %>%
  distinct() %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  filter(!is.na(`Hours Fished`), `Hours Fished` > 0) %>%
  group_by(Month) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE)
  )
df_table1 <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Vessels = n_distinct(`Name of Vessel`),
    Conch = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Conch / Hours, 1), NA)
  ) %>%
  bind_rows(df_table1_overall)

# Prepare dynamic variables for Table 1 ---------------------------
total_vessels <- length(unique(df$`Name of Vessel`))
total_conch <- df_table1_overall$Conch
total_events <- df_table1_overall$Missions
total_hours <- df_table1_overall$Hours
max_cpue_num <- max(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
max_cpue_num_month <- df_table1$Month[which.max(df_table1$`CPUE (num/hr)`)]
min_cpue_num <- min(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
min_cpue_num_month <- df_table1$Month[which.min(df_table1$`CPUE (num/hr)`)]
```


Between `r study_start` and `r study_end`, `r total_conch` conch were 
caught and recorded as part of the TAMR 
Fisheries Catch dataset. These conch were caught across `r total_events` missions, 
between `r total_vessels` surveyed vessels, 
totalling `r total_hours` active fishing hours.
The highest CPUE in count 
per hour (num/hr) occurred in `r max_cpue_num_month` 
with `r max_cpue_num`, whereas the lowest was `r min_cpue_num_month` 
in `r min_cpue_num` (Table 1; Figure 1).\

**Table 1.** Summary of fishing activity by month, including number of surveyed vessels, 
missions, and conch caught, along with fishing hours, pounds of catch, 
and catch per unit effort (CPUE). 
CPUE values are calculated only for missions where fishing hours and catch weight were recorded. 
Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\


```{r table1, echo = FALSE}
# show table
as.data.frame(df_table1)
```


<br>


```{r table1_expanded_prep, echo = FALSE, include = FALSE}
# Create df for Table 1 Expanded ---------------------------
df_table1_expanded_rows <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  group_by(`Mission Id`) %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  summarize(
    Month = first(Month),
    Month_sort = first(Month_sort),
    Hours = first(`Hours Fished`),
    Conch = n(),
    .groups = "drop"
  ) %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, Conch / Hours, NA)
  )
df_table1_expanded_num_overall <- df_table1_expanded_rows %>%
  summarize(
    Min = round(min(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (num/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (num/hr)`, na.rm = TRUE), 1)
  ) %>%
  mutate(Month = "Overall", CPUE = "num/hr")
df_table1expanded <- df_table1_expanded_rows %>%
  mutate(CPUE = "num/hr") %>%
  group_by(CPUE, Month, Month_sort) %>%
  summarize(
    Min = round(min(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Mean = round(mean(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Median = round(median(`CPUE (num/hr)`, na.rm = TRUE), 1),
    Max = round(max(`CPUE (num/hr)`, na.rm = TRUE), 1),
    SD = round(sd(`CPUE (num/hr)`, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  bind_rows(df_table1_expanded_num_overall)

# Prepare dynamic variables for Table 1 Expanded ---------------------------
overall_mean <- df_table1_expanded_num_overall$Mean
overall_median <- df_table1_expanded_num_overall$Median
overall_min <- df_table1_expanded_num_overall$Min
overall_max <- df_table1_expanded_num_overall$Max
overall_sd <- df_table1_expanded_num_overall$SD
max_cpue_mean_num <- max(df_table1expanded$Mean, na.rm = TRUE)
max_cpue_mean_num_month <- df_table1expanded$Month[which.max(df_table1expanded$Mean)]
min_cpue_mean_num <- min(df_table1expanded$Mean, na.rm = TRUE)
min_cpue_mean_num_month <- df_table1expanded$Month[which.min(df_table1expanded$Mean)]
```


The CPUE of conch in count per hour had a 
mean of `r overall_mean`num/hr and median of `r overall_median`num/hr, 
with the highest mean of `r max_cpue_mean_num`num/hr being in `r max_cpue_mean_num_month` 
and the lowest mean of `r min_cpue_mean_num`num/hr being in `r min_cpue_mean_num_month`. 
Minimum CPUE recorded across all missions was `r overall_min`num/hr, maximum was `r overall_max`num/hr, 
and standard deviation was `r overall_sd`num/hr 
(Table 1; Figure 1).\


```{r figure1_prep, echo = FALSE, include = FALSE}
# Create dfs for Figure 1 ---------------------------
df_fig1_A <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Vessels = n_distinct(`Name of Vessel`),
    Conch = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  mutate(
    `CPUE (num/hr)` = round(Conch / Hours, 1)
  )

# Create panels for Figure 1 ---------------------------
A <- ggplot(df_fig1_A, aes(x = Month_sort, y = Hours, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Hours Fished") +
  labs(title = "A")
B <- ggplot(df_fig1_A, aes(x = Month_sort, y = Conch, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Number Conch") +
  labs(title = "B")
C <- ggplot(df_fig1_A, aes(x = Month_sort, y = `CPUE (num/hr)`, group = 1)) +
  geom_line(size = 0.5) +
  geom_point(size = 2.5) +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  expand_limits(y = 0) +
  theme_custom +
  xlab("Month") +
  ylab("Catch per Unit Effort") +
  labs(title = "C")


# Assemble Figure 1 ---------------------------
figure1 <- ggarrange(
  A, B, C,
  nrow = 3, ncol = 1
)
```


```{r figure1, fig.width=6, fig.height=8, fig.align = 'left', echo = FALSE}
figure1
```


*Figure 1.* Line plots of 
active hours fished (A), conch caught (B), 
and catch per unit effort (in count per active hour fished) (C) 
over time. 
Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\




```{r table2_prep, echo = FALSE, include = FALSE}
# Create df for Table 2 ---------------------------
df_table2 <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01")),
    weight_for_mk = ifelse(`Weight Type` == "Market Clean", `Weight (g)`, NA_real_),
    weight_for_whole = ifelse(`Weight Type` == "Whole", `Weight (g)`, NA_real_),
    weight_for_up = ifelse(`Weight Type` == "Unprocessed", `Weight (g)`, NA_real_)
  ) %>%
  mutate(
    `Shell Length (mm)` = as.numeric(`Shell Length (mm)`),
    `Lip Thickness (mm)` = as.numeric(`Lip Thickness (mm)`),
    weight_for_mk = as.numeric(weight_for_mk),
    weight_for_whole = as.numeric(weight_for_whole),
    weight_for_up = as.numeric(weight_for_up)
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    `Mean Shell Length (mm)` = round(mean(`Shell Length (mm)`, na.rm = TRUE), 1),
    `Mean Lip Thickness (mm)` = round(mean(`Lip Thickness (mm)`, na.rm = TRUE), 1),
    `Mean Market Clean Weight (g)` = round(mean(weight_for_mk, na.rm = TRUE), 1),
    `Mean Whole Weight (g)` = round(mean(weight_for_whole, na.rm = TRUE), 1),
    `Mean Unprocessed Weight (g)` = round(mean(weight_for_up, na.rm = TRUE), 1)
  ) %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  bind_rows(
    df %>%
      mutate(
        weight_for_mk = ifelse(`Weight Type` == "Market Clean", `Weight (g)`, NA_real_),
        weight_for_whole = ifelse(`Weight Type` == "Whole", `Weight (g)`, NA_real_),
        weight_for_up = ifelse(`Weight Type` == "Unprocessed", `Weight (g)`, NA_real_)
      ) %>%
      mutate(
        `Shell Length (mm)` = as.numeric(`Shell Length (mm)`),
        `Lip Thickness (mm)` = as.numeric(`Lip Thickness (mm)`),
        weight_for_mk = as.numeric(weight_for_mk),
        weight_for_whole = as.numeric(weight_for_whole),
        weight_for_up = as.numeric(weight_for_up)
      ) %>%
      summarize(
        `Mean Shell Length (mm)` = round(mean(`Shell Length (mm)`, na.rm = TRUE), 1),
        `Mean Lip Thickness (mm)` = round(mean(`Lip Thickness (mm)`, na.rm = TRUE), 1),
        `Mean Market Clean Weight (g)` = round(mean(weight_for_mk, na.rm = TRUE), 1),
        `Mean Whole Weight (g)` = round(mean(weight_for_whole, na.rm = TRUE), 1),
        `Mean Unprocessed Weight (g)` = round(mean(weight_for_up, na.rm = TRUE), 1)
      ) %>%
      mutate(Month = "Overall")
  ) %>%
  mutate(across(everything(), ~ {
    if (is.numeric(.)) {
      ifelse(is.na(.) | is.nan(.), "-", as.character(.))
    } else {
      ifelse(is.na(.), "-", .)
    }
  }))

# Prepare dynamic variables for Table 2 ---------------------------
total_shell <- filter(df_table2, Month == "Overall")$`Mean Shell Length (mm)`
max_shell <- max(filter(df_table2, Month != "Overall")$`Mean Shell Length (mm)`, na.rm = TRUE)
max_shell_month <- df_table2$Month[which.max(filter(df_table2, Month != "Overall")$`Mean Shell Length (mm)`)]
total_lip <- filter(df_table2, Month == "Overall")$`Mean Lip Thickness (mm)`
max_lip <- max(filter(df_table2, Month != "Overall")$`Mean Lip Thickness (mm)`, na.rm = TRUE)
max_lip_month <- df_table2$Month[which.max(filter(df_table2, Month != "Overall")$`Mean Lip Thickness (mm)`)]
total_weight_mk <- filter(df_table2, Month == "Overall")$`Mean Market Clean Weight (g)`
max_weight_mk <- max(filter(df_table2, Month != "Overall")$`Mean Market Clean Weight (g)`, na.rm = TRUE)
max_weight_mk_month <- df_table2$Month[which.max(filter(df_table2, Month != "Overall")$`Mean Market Clean Weight (g)`)]
total_weight_whole <- filter(df_table2, Month == "Overall")$`Mean Whole Weight (g)`
max_weight_whole <- max(filter(df_table2, Month != "Overall")$`Mean Whole Weight (g)`, na.rm = TRUE)
max_weight_whole_month <- df_table2$Month[which.max(filter(df_table2, Month != "Overall")$`Mean Whole Weight (g)`)]
total_weight_up <- filter(df_table2, Month == "Overall")$`Mean Unprocessed Weight (g)`
max_weight_up <- max(filter(df_table2, Month != "Overall")$`Mean Unprocessed Weight (g)`, na.rm = TRUE)
max_weight_up_month <- df_table2$Month[which.max(filter(df_table2, Month != "Overall")$`Mean Unprocessed Weight (g)`)]
```

Average shell length was `r total_shell`oz, 
greatest in `r max_shell_month` with `r max_shell`oz. 
Average lip thickness was `r total_lip`oz, 
greatest in `r max_lip_month` with `r max_lip`oz. 
Average market clean weight was `r total_weight_mk`oz, 
greatest in `r max_weight_mk_month` with `r max_weight_mk`oz. 
Average whole weight was `r total_weight_whole`oz, 
greatest in `r max_weight_whole_month` with `r max_weight_whole`oz. 
Average unprocessed weight was `r total_weight_up`oz, 
greatest in `r max_weight_up_month` with `r max_weight_up`oz (Table 2; Figure 2; Figure 3).\


**Table 2.** Summary of mean shell length (mm), lip thickness (mm), market clean weight (oz), whole weight (oz), 
and unprocessed weight (oz) for fished conch. Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, 
`r study_timeframe`.\


```{r table2, echo = FALSE}
# show table
as.data.frame(df_table2)
```



```{r figure2_prep, echo = FALSE, include = FALSE}
# Create dfs for Figure 2 ---------------------------
df_fig2_unsummarized <- df %>%
  mutate(
    weight_for_mk = ifelse(`Weight Type` == "Market Clean", `Weight (g)`, NA_real_),
    weight_for_whole = ifelse(`Weight Type` == "Whole", `Weight (g)`, NA_real_),
    weight_for_up = ifelse(`Weight Type` == "Unprocessed", `Weight (g)`, NA_real_)
  ) %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01")),
    `Shell Length (mm)` = as.numeric(`Shell Length (mm)`),
    `Lip Thickness (mm)` = as.numeric(`Lip Thickness (mm)`),
    weight_for_mk = as.numeric(weight_for_mk),
    weight_for_whole = as.numeric(weight_for_whole),
    weight_for_up = as.numeric(weight_for_up)
  ) %>%
  arrange(Month_sort)
df_fig2 <- df %>%
  mutate(
    weight_for_mk = ifelse(`Weight Type` == "Market Clean", `Weight (g)`, NA_real_),
    weight_for_whole = ifelse(`Weight Type` == "Whole", `Weight (g)`, NA_real_),
    weight_for_up = ifelse(`Weight Type` == "Unprocessed", `Weight (g)`, NA_real_),
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01")),
    `Shell Length (mm)` = as.numeric(`Shell Length (mm)`),
    `Lip Thickness (mm)` = as.numeric(`Lip Thickness (mm)`)
  ) %>%
  mutate(
    weight_for_mk = as.numeric(weight_for_mk),
    weight_for_whole = as.numeric(weight_for_whole),
    weight_for_up = as.numeric(weight_for_up)
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    weight_for_mk = round(mean(weight_for_mk, na.rm = TRUE), 1),
    weight_for_whole = round(mean(weight_for_whole, na.rm = TRUE), 1),
    weight_for_up = round(mean(weight_for_up, na.rm = TRUE), 1),
    .groups = "drop"
  ) %>%
  arrange(Month_sort)

# Create panels for Figure 2 ---------------------------
A <- ggplot(df_fig2_unsummarized, aes(x = weight_for_mk)) +
  geom_histogram(binwidth = 10, color = "black") +
  scale_fill_manual(values = palette[2:3], drop = FALSE) +
  theme_custom +
  expand_limits(x = 0) +
  xlab("Market Clean Weight (g)") +
  ylab("Count") +
  labs(title = "A")
B <- ggplot(df_fig2_unsummarized, aes(x = weight_for_whole)) +
  geom_histogram(binwidth = 100, color = "black") +
  scale_fill_manual(values = palette[2:3], drop = FALSE) +
  theme_custom +
  expand_limits(x = 0) +
  xlab("Whole Weight (g)") +
  ylab("Count") +
  labs(title = "B")
C <- ggplot(df_fig2, aes(x = Month_sort)) +
  theme_custom +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  ylab("Weight (g)") +
  xlab("Month") +
  scale_shape_manual(
    name = "Weight Type",
    values = c("Market Clean" = 4, "Whole" = 19, "Unprocessed" = 6)
  ) +
  scale_color_manual(
    name = "Weight Type",
    values = c("Market Clean" = "black", "Whole" = "black", "Unprocessed" = "black")
  ) +
  expand_limits(y = 0) +
  labs(title = "C") +
  theme(legend.position = "bottom")

# Add points and lines per weight type ------------------------------
if (sum(!is.na(df_fig2$weight_for_mk)) >= 2) {
  C <- C + geom_line(aes(y = weight_for_mk, group = 1), size = 0.5)
}
if (sum(!is.na(df_fig2$weight_for_mk)) >= 1) {
  C <- C + geom_point(aes(y = weight_for_mk, shape = "Market Clean"), size = 2.5)
}
if (sum(!is.na(df_fig2$weight_for_whole)) >= 2) {
  C <- C + geom_line(aes(y = weight_for_whole, group = 1), size = 0.5)
}
if (sum(!is.na(df_fig2$weight_for_whole)) >= 1) {
  C <- C + geom_point(aes(y = weight_for_whole, shape = "Whole"), size = 2.5)
}
if (sum(!is.na(df_fig2$weight_for_up)) >= 2) {
  C <- C + geom_line(aes(y = weight_for_up, group = 1), size = 0.5)
}
if (sum(!is.na(df_fig2$weight_for_up)) >= 1) {
  C <- C + geom_point(aes(y = weight_for_up, shape = "Unprocessed"), size = 2.5)
}


# Assemble Figure 2 ---------------------------
figure_3_top <- ggarrange(
  A, B,
  ncol = 2, nrow = 1
)
figure2 <- ggarrange(
  figure_3_top, C,
  ncol = 1, nrow = 2
)
```


```{r figure2, fig.width=6, fig.height=6, fig.align = 'left', echo = FALSE}
figure2
```


*Figure 2.* Histogram of market clean weight (g) (A), histogram of whole weight (g) (B), and 
line plot of different average weight types recorded over time (C).
Data were collected using Fisheries catch lobster data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\


```{r figure3_prep, echo = FALSE, include = FALSE}
# Create df for Figure 3 ---------------------------
df_fig3 <- df %>%
  mutate(
    `Shell Length (cm)` = as.numeric(`Shell Length (mm)`) / 10,
    `Lip Thickness (mm)` = as.numeric(`Lip Thickness (mm)`)
  )

# Create Figure 3 ---------------------------
A <- ggplot(filter(df_fig3, is.na(`Lip Thickness (mm)`)), aes(x = `Shell Length (cm)`)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5)) +
  labs(title = "A", subtitle = "No Data for Lip Thickness")
B <- ggplot(filter(df_fig3, `Lip Thickness (mm)` <= 5 & `Lip Thickness (mm)` >= 0), aes(x = `Shell Length (cm)`)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5)) +
  labs(title = "B", subtitle = "Lip Thickness 0-5mm")
C <- ggplot(filter(df_fig3, `Lip Thickness (mm)` > 5), aes(x = `Shell Length (cm)`)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5)) +
  labs(title = "C", subtitle = "Lip Thickness >5mm")
D <- ggplot(df_fig3, aes(x = `Shell Length (cm)`)) +
  geom_histogram(color = "black", position = "identity", alpha = 0.5) +
  theme_custom +
  scale_fill_manual(values = c(palette[3], palette[4])) +
  geom_vline(xintercept = 17.8, linetype = "dashed") +
  scale_x_continuous(limits = c(0, 30), breaks = seq(0, 30, 5)) +
  labs(title = "D", subtitle = "Overall")

# Assemble Figure 3 ---------------------------
figure3 <- ggarrange(A, B, C, D, nrow = 2, ncol = 2)
```


```{r figure3, fig.width=6, fig.height=5, fig.align = 'left', echo = FALSE}
figure3
```


*Figure 3.* Length-frequency distribution of lipped and non-lipped queen conch shell length (cm) at Turneffe Atoll Marine Reserve,
for conch with no lip thickness data (A), conch with a lip from 0-5mm (B), conch with a lip >5mm (C), and all recordered conch (D).  
The dotted vertical lines indicate the legal shell length (17.8 cm) for harvest. Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, 
`r study_timeframe`.\


```{r table3, echo = FALSE, include = FALSE}
# Create df for Table 3 ---------------------------
df_table3_separate <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  select(`Mission Id`, Month, Month_sort, `Vessel Type`) %>%
  group_by(`Vessel Type`, Month) %>%
  summarize(
    Vessels = n(),
    .groups = "drop"
  ) %>%
  mutate(Month = as.character(Month))
df_table3_overall <- df %>%
  mutate(
    Month = "Overall",
    Month_sort = "Overall"
  ) %>%
  select(`Mission Id`, Month, Month_sort, `Vessel Type`) %>%
  group_by(`Vessel Type`, Month) %>%
  summarize(
    Vessels = n(),
    .groups = "drop"
  )
df_table3 <- bind_rows(df_table3_separate, df_table3_overall)
df_table3 <- df_table3 %>%
  mutate(`Vessel Type` = case_when(
    `Vessel Type` %in% c("Canoe", "canoe", "Sailboat") ~ "Canoe/Sailboat",
    TRUE ~ `Vessel Type`
  )) %>%
  pivot_wider(names_from = `Vessel Type`, values_from = `Vessels`, values_fn = sum, values_fill = 0)
df_table3_notoverall <- filter(df_table3, Month != "Overall")

# Prepare dynamic variables for Table 3 ---------------------------
max_skiff <- max(df_table3_notoverall$Skiff, na.rm = TRUE)
max_skiff_month <- df_table3_notoverall$Month[which.max(df_table3_notoverall$Skiff)]
max_canoe <- max(df_table3_notoverall$`Canoe/Sailboat`, na.rm = TRUE)
max_canoe_month <- df_table3_notoverall$Month[which.max(df_table3_notoverall$`Canoe/Sailboat`)]
```


Throughout the study period, fishers used multiple vessel types to complete missions. 
Skiffs were used the most in `r max_skiff_month` with `r max_skiff` instances, 
wheres sailboats/canoes were used the most in `r max_canoe_month` with `r max_canoe` instances (Table 3; Figure 4).\


**Table 3.** Count of vessel type counts for each month, where one count corresponds to a vessel submitting data for a single mission. 
Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\


```{r df_table3, echo = FALSE}
# show table
as.data.frame(df_table3)
```



```{r figure4_prep, echo = FALSE, include = FALSE}
# Create dfs for Figure 4 ---------------------------
df_fig4 <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  select(`Mission Id`, Month, Month_sort, `Vessel Type`) %>%
  mutate(`Vessel Type` = case_when(
    `Vessel Type` %in% c("Canoe", "canoe", "Sailboat") ~ "Canoe/Sailboat",
    TRUE ~ `Vessel Type`
  )) %>%
  group_by(`Vessel Type`, Month_sort) %>%
  summarize(
    Vessels = n(),
    .groups = "drop"
  )

# Create Figure 4 ---------------------------
figure4 <- ggplot(df_fig4, aes(x = Month_sort, color = `Vessel Type`, y = Vessels)) +
  geom_line() +
  geom_point(color = "black") +
  scale_color_manual(values = palette[2:5]) +
  theme_custom +
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b '%y"
  ) +
  xlab("Month") +
  ylab("Number of Vessels")
```


```{r figure4, fig.width=6, fig.height=2.5, fig.align = 'left', echo = FALSE}
figure4
```


*Figure 4.* Line and scatter plot of vessel type counts for each month, where one count corresponds to a vessel submitting data for a single mission. 
Data were collected using Fisheries catch conch data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\



<br>


## Conclusion

By summarizing and visualizing the fisheries catch conch data from `r study_timeframe`, 
researchers, policy-makers, and other stakeholders are better able to interpret and use the dataset. 
This report supports the aims of the Fisheries Catch data collection by offering a more understandable view 
of the data itself, supporting future efforts to create recommendations on enforcement, monitoring, 
education, outreach, and regulations to ensure the long-term sustainability of the TAMR fishery while safeguarding the 
ecological value of the atoll.\


## Acknowledgments

Thank you to the data collectors and fishers who contributed to this project.\


## Report Metadata

**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\
**App Version:** `r params$release`\


## Literature Cited

Appeldoorn, R. S., & Baker, N. (2013). A Literature Review of the Queen Conch, Strombus gigas [Technical Report]. University of Puerto Rico.

Belize Fisheries Department. (n.d.). Capture Fisheries Unit. https://fisheries.gov.bz/units/cfu/

Fedler, A. J. (2011). The Economic Value of Turneffe Atoll (Full Report). Turneffe Atoll Trust. http://www.turneffeatollmarinereserve.org/app/webroot/userfiles/214/File/Turneffe%20Atoll%20Valuation%20Study%20PDF%20-%20Final%20Report.pdf

Meerman, J. C. (2013). Belize Basemap [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Meerman, J. C. (2017). Belize_Ecosystems_2017 [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

NOAA Fisheries. (2025). Queen Conch: Science. https://www.fisheries.noaa.gov/species/queen-conch/science

Stoddart, D. R. (1962). Three Caribbean atolls: Turneffe Islands, Lighthouse Reef and Glover’s Reef,  British Honduras. Atoll Research Bulletin, 87, 1–151.

Turneffe Atoll Sustainability Association. (2025). Turneffe Reef-Monitoring Data Reporting Tool (Version 1.0) [Computer software].
