---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  release: "Unknown"
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  fisheries_1per_timeframe: "Annual"
  fisheries_1per_year_selection: "None"
  fisheries_1per_period_selection: "None"
  fisheries_1per_lobster_season_selection: "None" 
  fisheries_1per_conch_season_selection: "None" 
  fisheries_1per_finfish_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Determine year or season  ---------------------------
timeframe <- params$fisheries_1per_timeframe

# Filter to year or season  ---------------------------
df <- params$datafile[[1]]
df$`Date of Encounter` <- as.Date(df$`Date of Encounter`)
if (timeframe == "Annual" && as.character(params$fisheries_1per_year_selection) != "None") {
  df <- df |>
    filter(format(`Date of Encounter`, "%Y") == params$fisheries_1per_year_selection)
} else if (timeframe == "Seasonal" && as.character(params$fisheries_1per_conch_season_selection) != "None") {
  season_years <- unlist(strsplit(params$fisheries_1per_conch_season_selection, "-"))
  start_date <- as.Date(paste0(season_years[1], "-07-01"))
  end_date <- as.Date(paste0(season_years[2], "-02-28"))
  df <- df |>
    filter(`Date of Encounter` >= start_date, `Date of Encounter` <= end_date)
}

# Add month column  ---------------------------
df$Month <- format(df$`Date of Encounter`, "%b %Y")

# Define study timeframe ---------------------------
study_start <- min(df$`Date of Encounter`, na.rm = TRUE)
study_end <- max(df$`Date of Encounter`, na.rm = TRUE)
if (timeframe == "Annual") {
  study_timeframe <- unique(format(df$`Date of Encounter`, "%Y"))[1]
} else if (timeframe == "Seasonal") {
  study_timeframe <- paste0(format(study_start, "%Y"), "-", format(study_end, "%Y"))
}
```

# Turneffe Atoll Fisheries Catch Finfish `r timeframe` Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)


**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\
**App Version:** `r params$release`\

***

Coral reef ecosystems, like those of Turneffe Atoll Marine Reserve (TAMR), provide key ecological and economic value to coastal families. 
With proper management and monitoring, 
these ecosystems can continue to support people sustainably in the long-term. 
Directly tracking the activities and harvests of local fishers is a key method for monitoring coral reefs, 
as it offers insights into ecosystem dynamics and possible pressures. 
This report presents `r tolower(timeframe)` findings from fishery-dependent monitoring of finfish 
at Turneffe Atoll Marine Reserve (TAMR) from `r study_timeframe`. 
Data were recorded by TASA patrolling officers in direct communication with TAMR fishers during their fishing trips. 
Data were collected on key economic species. 
This report aims to summarize and visualize the Fisheries 
Finfish Catch dataset from `r study_timeframe` to facilitate data interpretation and strengthen the scientific basis for 
management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Fisheries Catch Project

### Finfish

Finfish refer to all fish species that can be classified as ‘true fish,’ in contrast to other aquatic animals 
such as shellfish or cuttlefish.
In Belize, over 650 species of finfish may be found (FishBase 2024), with many of these, such as 
Mutton Snapper (_Lutjanus analis_),
Nassau Grouper (_Epinephelus striatus_), and Queen Triggerfish (_Balistes vetula_) being 
listed as Vulnerable conservation status on the IUCN.

Finfish provide a variety of services to coral reefs, including grazing of macroalgae and predation.
Fishing for finfish may be done using handline fishing or winch (targeting snappers, jacks, groupers, and grunts primarily),
long lines (targeting pelagic species and sharks), or gillnets (targeting sharks) in Belize (UNCTAD 2020).
Finfish fishing in Belize is regulated by the Belize Fisheries Department (BFD),
with both general regulations such as fishers requiring fishing licenses and color-coded vessels, and specific regulations,
such as forbidding the fishing of parrotfish and tangs (Scaridae) (Belize Fisheries Department accessed 2025-b).
To maintain healthy stocks and ecological services of finfish, it is important to maintain the size, reproductive capacity,
and biodiversity of key taxa.\


## Methods

This report, including text, figures, and tables, was generated using the 
Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025, version `r params$release`), an R Shiny app which takes 
standardized reef monitoring data, validates the dataset, and creates a shareable output. \

 
## Results


```{r table1_prep, echo = FALSE, include = FALSE}
# Create df for Table 1 ---------------------------
df_events_overall <- df %>%
  select(`Mission Id`, `Hours Fished`) %>%
  distinct() %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  filter(!is.na(`Hours Fished`), `Hours Fished` > 0) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE)
  ) %>%
  mutate(Month = "Overall")
df_table1_overall <- df %>%
  summarize(
    Vessels = n_distinct(`Name of vessel used`),
    Finfish = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  mutate(Month = "Overall") %>%
  left_join(df_events_overall, by = "Month") %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  )
df_events <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  select(Month, `Mission Id`, `Hours Fished`) %>%
  distinct() %>%
  mutate(`Hours Fished` = as.numeric(`Hours Fished`)) %>%
  filter(!is.na(`Hours Fished`), `Hours Fished` > 0) %>%
  group_by(Month) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE)
  )
df_table1 <- df %>%
  mutate(
    Month = format(`Date of Encounter`, "%b %Y"),
    Month_sort = as.Date(format(`Date of Encounter`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Vessels = n_distinct(`Name of vessel used`),
    Finfish = n(),
    Missions = n_distinct(`Mission Id`)
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  mutate(
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  ) %>%
  bind_rows(df_table1_overall)

# Prepare dynamic variables for Table 1 ---------------------------
total_vessels <- length(unique(df$`Name of vessel used`))
total_finfish <- df_table1_overall$Finfish
total_events <- df_table1_overall$Missions
total_hours <- df_table1_overall$Hours
max_cpue_num <- max(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
max_cpue_num_month <- df_table1$Month[which.max(df_table1$`CPUE (num/hr)`)]
min_cpue_num <- min(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
min_cpue_num_month <- df_table1$Month[which.min(df_table1$`CPUE (num/hr)`)]
```


Between `r study_start` and `r study_end`, `r total_finfish` Caribbean spiny finfish were 
caught and recorded as part of the TAMR 
Fisheries Catch dataset. These finfish were caught across `r total_events` missions, 
between `r total_vessels` surveyed vessels, 
totalling `r total_hours` active fishing hours.
The highest CPUE in count 
per hour (num/hr) occurred in `r max_cpue_num_month` 
with `r max_cpue_num`, whereas the lowest was `r min_cpue_num_month` 
in `r min_cpue_num` (Table 1; Figure 1).\

**Table 1.** Summary of fishing activity by month, including number of surveyed vessels, 
missions, and finfish caught, along with fishing hours, pounds of catch, 
and catch per unit effort (CPUE). 
CPUE values are calculated only for missions where fishing hours and catch weight were recorded. 
Data were collected using Fisheries catch finfish data from Turneffe Atoll Marine Reserve, `r study_timeframe`.\


```{r table1, echo = FALSE}
# show table
as.data.frame(df_table1)
```


<br>
