---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  fisher_1per_timeframe: "Annual"
  fisher_1per_year_selection: "None"
  fisher_1per_period_selection: "None"
  fisher_1per_lobster_season_selection: "None" 
  fisher_1per_conch_season_selection: "None" 
  fisher_1per_finfish_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Determine year or season  ---------------------------
timeframe <- "Annual"

# Filter to year or season  ---------------------------
df <- params$datafile[[1]]
df$`Waypoint Date` <- as.Date(df$`Waypoint Date`)
if (timeframe == "Annual" && as.character(params$fisher_1per_year_selection) != "None") {
  df <- df |>
    filter(format(`Waypoint Date`, "%Y") == params$fisher_1per_year_selection)
}

# Add month column  ---------------------------
df$Month <- format(df$`Waypoint Date`, "%b %Y")

# Define study timeframe ---------------------------
study_start <- min(df$`Waypoint Date`, na.rm = TRUE)
study_end <- max(df$`Waypoint Date`, na.rm = TRUE)
study_timeframe <- unique(format(df$`Waypoint Date`, "%Y"))[1]
```

# Turneffe Atoll Fisher Catch Finfish `r timeframe` Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)


**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\

***


The marine resources and biodiversity abundant in coral reef ecosystems support the 
livelihoods of coastal communities, and provide economic opportunities that can be 
sustainable given attentive management and enforcement. 
Monitoring the activities and harvests of local fishers is important for 
understanding ecosystem productivity, assessing stock sustainability, 
and informing effective management decisions. 
This report presents `r tolower(timeframe)` findings from fishery-dependent monitoring of finfish 
at Turneffe Atoll Marine Reserve (TAMR) from `r study_timeframe`. 
Data were recorded directly by fishermen using SMART technology through a data collection 
package designed by TASA. This report aims to summarize and visualize the Fisher 
Finfish Catch dataset from `r study_timeframe` to facilitate data interpretation and support 
management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Fisher Catch Project

Background on reef monitoring dataset type \


### Finfish

Finfish refer to all fish species that can be classified as ‘true fish,’ in contrast to other aquatic animals 
such as shellfish or cuttlefish.
In Belize, over 650 species of finfish may be found (FishBase 2024), with many of these, such as 
Mutton Snapper (_Lutjanus analis_),
Nassau Grouper (_Epinephelus striatus_), and Queen Triggerfish (_Balistes vetula_) being 
listed as Vulnerable conservation status on the IUCN.

Finfish provide a variety of services to coral reefs, including grazing of macroalgae and predation.
Fishing for finfish may be done using handline fishing or winch (targeting snappers, jacks, groupers, and grunts primarily),
long lines (targeting pelagic species and sharks), or gillnets (targeting sharks) in Belize (UNCTAD 2020).
Finfish fishing in Belize is regulated by the Belize Fisheries Department (BFD),
with both general regulations such as fishers requiring fishing licenses and color-coded vessels, and specific regulations,
such as forbidding the fishing of parrotfish and tangs (Scaridae) (Belize Fisheries Department accessed 2025-b).
To maintain healthy stocks and ecological services of finfish, it is important to maintain the size, reproductive capacity,
and biodiversity of key taxa.\


## Methods

Fisher catch data were collected at Turneffe Atoll 
Marine Reserve (TAMR) (Figure 1). 
The survey period for this report’s dataset extended 
from `r study_start` to `r study_end` at TAMR.

Fishermen were selected [process of selecting fishermen 
for the project]

Data on gutted or whole weight (g), total length (cm), and fork length (cm) were entered by fishermen using SMART 
technology through a data collection package 
designed by TASA. They [process of entering data by 
fishermen into SMART]

This report, including text, figures, and tables, was generated using the 
Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025), an R Shiny app which takes 
standardized reef monitoring data, validates the dataset, and creates a shareable output. \


```{r figure1, echo = FALSE, fig.width = 7, fig.height = 7}
# Create map for Figure 1 ---------------------------
df_map <- df |>
  select(`Waypoint ID`, Y, X) |>
  distinct()
if (length(unique(df$`Waypoint ID`)) < 5) {
  generate_map(df_map, "Y", "X", "Waypoint ID")
} else {
  generate_map_nolabels(df_map, "Y", "X", "Waypoint ID")
}
```


*Figure 1.* Map of Turneffe Atoll Marine Reserve zones and ecosystems. 
Shapefiles for zones belong to the Belize Fisheries Department (Belize Geomatics 2022). 
Shapefiles for the base map and ecosystems come from the publicly available Spatial Data 
Warehouse for Belize (Meerman 2013, Meerman 2017).
Waypoints for finfish recorded in `r study_timeframe` are labelled on the map.\

 
## Results


```{r table1_prep, echo = FALSE, include = FALSE}
df_events_overall <- df %>%
  select(`Waypoint ID`, `Hours Fished`, `Total Lbs of Catch`) %>%
  distinct() %>%
  filter(!is.na(`Hours Fished`), !is.na(`Total Lbs of Catch`), `Hours Fished` > 0) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE),
    `Catch (lb)` = sum(`Total Lbs of Catch`, na.rm = TRUE)
  ) %>%
  mutate(Month = "Overall")
df_table1_overall <- df %>%
  summarize(
    Fishers = n_distinct(`Fishers Name`),
    Finfish = n(),
    Events = n_distinct(`Waypoint ID`)
  ) %>%
  mutate(Month = "Overall") %>%
  left_join(df_events_overall, by = "Month") %>%
  mutate(
    `CPUE (lb/hr)` = ifelse(Hours > 0, round(`Catch (lb)` / Hours, 1), NA),
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  )
df_events <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  select(Month, `Waypoint ID`, `Hours Fished`, `Total Lbs of Catch`) %>%
  distinct() %>%
  filter(!is.na(`Hours Fished`), !is.na(`Total Lbs of Catch`), `Hours Fished` > 0) %>%
  group_by(Month) %>%
  summarize(
    Hours = sum(`Hours Fished`, na.rm = TRUE),
    `Catch (lb)` = sum(`Total Lbs of Catch`, na.rm = TRUE)
  )
df_table1 <- df %>%
  mutate(
    Month = format(`Waypoint Date`, "%b %Y"),
    Month_sort = as.Date(format(`Waypoint Date`, "%Y-%m-01"))
  ) %>%
  group_by(Month, Month_sort) %>%
  summarize(
    Fishers = n_distinct(`Fishers Name`),
    Finfish = n(),
    Events = n_distinct(`Waypoint ID`)
  ) %>%
  left_join(df_events, by = "Month") %>%
  arrange(Month_sort) %>%
  select(-Month_sort) %>%
  mutate(
    `CPUE (lb/hr)` = ifelse(Hours > 0, round(`Catch (lb)` / Hours, 1), NA),
    `CPUE (num/hr)` = ifelse(Hours > 0, round(Finfish / Hours, 1), NA)
  ) %>%
  bind_rows(df_table1_overall)

# Prepare dynamic variables for Table 1 ---------------------------
total_fishers <- length(unique(df$`Fishers Name`))
total_finfish <- df_table1_overall$Finfish
total_events <- df_table1_overall$Events
total_hours <- df_table1_overall$Hours
max_cpue_lb <- max(df_table1$`CPUE (lb/hr)`, na.rm = TRUE)
max_cpue_lb_month <- df_table1$Month[which.max(df_table1$`CPUE (lb/hr)`)]
min_cpue_lb <- min(df_table1$`CPUE (lb/hr)`, na.rm = TRUE)
min_cpue_lb_month <- df_table1$Month[which.min(df_table1$`CPUE (lb/hr)`)]
max_cpue_num <- max(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
max_cpue_num_month <- df_table1$Month[which.max(df_table1$`CPUE (num/hr)`)]
min_cpue_num <- min(df_table1$`CPUE (num/hr)`, na.rm = TRUE)
min_cpue_num_month <- df_table1$Month[which.min(df_table1$`CPUE (num/hr)`)]
```


Between `r study_start` and `r study_end`, `r total_fishers` fishermen contributed data to the queen finfish SMART dataset. 
A total of `r total_finfish` finfish were caught across `r total_events` 
fishing events, totalling `r total_hours` active fishing hours. 
The highest catch per unit effort (CPUE) 
in pounds per hour (lb/hr) occurred in `r max_cpue_lb_month` with `r max_cpue_lb`, whereas the lowest 
occurred in `r min_cpue_lb_month` with `r min_cpue_lb`. The highest CPUE in count 
per hour (num/hr) occurred in `r max_cpue_num_month` 
with `r max_cpue_num`, whereas the lowest was `r min_cpue_num_month` 
in `r min_cpue_num` (Table 1; Figure 2). 
CPUE values were calculated only using fishing events that had 
values for Hours Fished and Total Lbs of Catch, to ensure CPUE is representative.\


**Table 1.** Summary of fishing activity by month, including total fishers, events, and finfish caught, 
along with fishing hours, pounds of catch, and catch per unit effort (CPUE). 
CPUE values are calculated only for events where both fishing hours and catch weight were recorded. 
Data were collected using SMART finfish data from fishers at Turneffe Atoll Marine Reserve, `r study_timeframe`.\

```{r table1, echo = FALSE}
# show table
as.data.frame(df_table1)
```


<br>


## Conclusion

By summarizing and visualizing the fisher catch finfish SMART data from `r study_timeframe`, 
researchers, policy-makers, and other stakeholders are better able to interpret and use the dataset. 
This report supports the aims of the Fisher Catch SMART data collection by offering a more understandable view 
of the data itself, supporting future efforts to create recommendations on enforcement, monitoring, 
education, outreach, and regulations to ensure the long-term sustainability of the TAMR fishery while safeguarding the 
ecological value of the atoll.\


## Acknowledgments

Thank you to the fishers who contributed to this project.\


## Literature Cited

Belize Fisheries Department. (n.d.). Fisheries Regulations. https://fisheries.gov.bz/regulations/

Fedler, A. J. (2011). The Economic Value of Turneffe Atoll (Full Report). Turneffe Atoll Trust. http://www.turneffeatollmarinereserve.org/app/webroot/userfiles/214/File/Turneffe%20Atoll%20Valuation%20Study%20PDF%20-%20Final%20Report.pdf

Froese, R., & Pauly, D. (2024). FishBase [World Wide Web electronic publication.]. www.fishbase.org

Meerman, J. C. (2013). Belize Basemap [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Meerman, J. C. (2017). Belize_Ecosystems_2017 [Dataset]. Biodiversity and Environmental Resource Data System of Belize. http://www.biodiversity.bz/

Stoddart, D. R. (1962). Three Caribbean atolls: Turneffe Islands, Lighthouse Reef and Glover’s Reef,  British Honduras. Atoll Research Bulletin, 87, 1–151.

Turneffe Atoll Sustainability Association. (2025). Turneffe Reef-Monitoring Data Reporting Tool (Version 1.0) [Computer software].

United Nations Conference (UNCTAD) & on Trade and Development. (2020). Oceans Economy and Trade Strategy: Belize Marine Fisheries and Seafood Processing (UNCTAD/DITC/TED/INF/2020/5). https://unctad.org/system/files/official-document/ditctedinf2020d5_en.pdf