---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  release: "Unknown"
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  spag_1per_spag_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Filter to season
df <- params$datafile[[1]]
df$Date <- as.Date(df$Date)
if (as.character(params$spag_1per_spag_season_selection) != "None") {
  season_years <- unlist(strsplit(params$spag_1per_spag_season_selection, "-"))
  start_date <- as.Date(paste0(season_years[1], "-12-01"))
  end_date <- as.Date(paste0(season_years[2], "-05-31"))
  df <- df |>
    filter(Date >= start_date, Date <= end_date)
}
# Add month column  ---------------------------
df$Month <- format(df$Date, "%b %Y")

# Define study timeframe ---------------------------
study_start <- min(df$Date, na.rm = TRUE)
study_end <- max(df$Date, na.rm = TRUE)
study_timeframe <- paste0(format(study_start, "%Y"), "-", format(study_end, "%Y"))
```

# Turneffe Atoll Spawning Aggregations Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)


**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\
**App Version:** `r params$release`\

***

SUMMARY GOES HERE

## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Spawning Aggregations Monitoring

INFO HERE


## Methods

SPAGs data were collected at Turneffe Atoll 
Marine Reserve (TAMR) (Figure 1). 
The survey period for this report’s dataset extended 
from `r study_start` to `r study_end` at TAMR.


```{r figure1, echo = FALSE, fig.width = 7, fig.height = 7}
# Create map for Figure 1 ---------------------------
df_map <- df |>
  select(Site, Latitude, Longitude) |>
  distinct() |>
  na.omit()
if (study_year >= 2024) {
generate_map(df_map, "Latitude", "Longitude", "Site")
} else{
  generate_map_historic(df_map, "Latitude", "Longitude", "Site")
}
```

*Figure 1.* Map of Turneffe Atoll Marine Reserve zones and ecosystems. 
Shapefiles for zones belong to the Belize Fisheries Department (Belize Geomatics 2022). 
Shapefiles for the base map and ecosystems come from the publicly available Spatial Data 
Warehouse for Belize (Meerman 2013, Meerman 2017).
Waypoints for SPAG dives recorded in `r study_timeframe` are labelled on the map.\

Dive surveys were conducted by trained dive teams that followed standardized survey protocol at 
each Turneffe Atoll SPAG site. Dive timings were chosen to align with 
anticipated spawning events, which are associated with the full moon. 
For each dive, the GPS coordinates of the dive, start and end time of the dive, 
surface conditions, and underwater conditions were recorded. 
Next, species counts were taken, with observed fish species being counted and categorized into 
standardized size classes where it was possible to do so accurately. Otherwise, 
raw counts were simply taken without size differentiation. 
Behaviors indicative of spawning activity, including grouping, fighting, 
color changes, bite wounds, gravid individuals, courtship, and observed spawning, were also recorded. 

During the visual counts, Nassau groupers and other select fish species were additionally measured 
using laser sizing methodology, to acquire a more accurate measurement of length. The laser 
system comprised of two waterproof, dive-rated red lasers which were attached to 
a metal camera mounting frame. The lasers were configured to be set parallel to each other and aligned, 
with an exactly 12cm separation and a camera affixed to the top of the frame. 
During laser sizing, the lasers followed the 
lateral line of the fish as they swam by at approximately 5 to 6 meters away, 
and their movements were recorded via video. After the fieldwork, the videos were 
reviewed and the laser points were used to establish the scale for extrapolating the total lengths 
of the target fish. 

For a complete, detailed description of the methodology, please refer to the 
Reef Fish Spawning Aggregation Monitoring Protocol for the Meso-American Reef and the Wider Caribbean, 
Version 2.0 (2004).

This report, including text, figures, and tables, was generated using the 
Turneffe Reef-Monitoring Data Reporting Tool (TASA 2025, version `r params$release`), an R Shiny app which takes 
standardized reef monitoring data, validates the dataset, and creates a shareable output. \

## Results


## Conclusion

CONCLUSION


## Acknowledgments



## Literature Cited

Reef Fish Spawning Aggregation Monitoring Protocol for the Meso-American Reef and the Wider Caribbean, Version 2.0. (2004). https://reefresilience.org/pdf/Spawning_Aggregation_Monitoring_Protocol_4July2004.pdf
