---
output:
  officedown::rdocx_document:
    reference_docx: report_template.docx
    tables:
      style: table_template
params:
  user_name: "Unknown"
  datafile: pressure
  datafile_name: "Unknown"
  fisher_1per_timeframe: "Annual"
  fisher_1per_year_selection: "None"
  fisher_1per_period_selection: "None"
  fisher_1per_lobster_season_selection: "None" 
  fisher_1per_conch_season_selection: "None" 
  fisher_1per_finfish_season_selection: "None"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

# Load packages ---------------------------
library(knitr)
library(officedown)
library(rmarkdown)
library(tidyverse)
library(ggpubr)

# Source code  ---------------------------
source("theme.r")
source("map.r")

# Determine year or season  ---------------------------
timeframe <- params$fisher_1per_timeframe

# Filter to year or season  ---------------------------
df <- params$datafile[[1]]
df$`Waypoint Date` <- as.Date(df$`Waypoint Date`)
if (timeframe == "Annual" && as.character(params$fisher_1per_year_selection) != "None") {
  df <- df |>
    filter(format(`Waypoint Date`, "%Y") == params$fisher_1per_year_selection)
} else if (timeframe == "Seasonal" && as.character(params$fisher_1per_lobster_season_selection) != "None") {
  season_years <- unlist(strsplit(params$fisher_1per_lobster_season_selection, "-"))
  start_date <- as.Date(paste0(season_years[1], "-07-01"))
  end_date <- as.Date(paste0(season_years[2], "-02-28"))
  df <- df |>
    filter(`Waypoint Date` >= start_date, `Waypoint Date` <= end_date)
}

# Define study timeframe ---------------------------
study_start <- min(df$`Waypoint Date`, na.rm = TRUE)
study_end <- max(df$`Waypoint Date`, na.rm = TRUE)
if (timeframe == "Annual") {
  study_timeframe <- unique(format(df$`Waypoint Date`, "%Y"))[1]
} else if (timeframe == "Seasonal") {
  study_timeframe <- paste0(format(study_start, "%Y"), "-", format(study_end, "%Y"))
}
```

# Turneffe Atoll Fisher Catch Lobster `r timeframe` Dataset Report: `r study_timeframe`


Turneffe Atoll Sustainability Association (TASA)


**Date Generated:** `r format(Sys.time(), '%Y-%m-%d')` \
**Source Data:** `r params$datafile_name` \
**Generated by:** `r params$user_name`\

***


The marine resources and biodiversity abundant in coral reef ecosystems support the 
livelihoods of coastal communities, and provide economic opportunities that can be 
sustainable given attentive management and enforcement. 
Monitoring the activities and harvests of local fishers is important for 
understanding ecosystem productivity, assessing stock sustainability, 
and informing effective management decisions. 
This report presents `r tolower(timeframe)` findings from fishery-dependent monitoring of Caribbean spiny 
lobster (Panulirus argus) at Turneffe Atoll Marine Reserve (TAMR) from `r study_timeframe`. 
Data were recorded directly by fishermen using SMART technology through a data collection 
package designed by TASA. This report aims to summarize and visualize the Fisher 
Lobster Catch dataset from `r study_timeframe` to facilitate data interpretation and support 
management decisions by TASA staff, collaborators, and stakeholders.\


## Background

### Turneffe Atoll Marine Reserve

Turneffe Atoll is the largest atoll in Belize, spanning mangrove and littoral forest cayes, seagrass beds, and coral reef ecosystems (Stoddart, 1962). 
It features an inner lagoon as well as an outer reef that encircles the atoll. 
The Turneffe Atoll Marine Reserve (TAMR), which encompasses the entire atoll, covers 1,471 km² and includes multiple management zones, such as general use, conservation, preservation, and spawning aggregation zones, as well as special management areas. 
TAMR plays a valuable role in supporting biodiversity and sustaining the communities that depend on its resources. 
Its reef provides coastal protection from extreme weather events (Fedler, 2011), serves as a habitat for diverse marine species, and supports economic activities such as commercial and recreational fishing, guided diving, and tourism.\

### Fisher Catch Project

Background on reef monitoring dataset type \

Map of study area \

## Methods

Survey methods\ 

 
**Table 1.** Table of surveys/dives.

_Dive Surveys_
```{r table1, echo = FALSE}
# show table
head(df)
```
